---
title: "Tier 3 Report - Smallmouth"
author: "Aaron Dain"
always_allow_html: true
output:
  bookdown::html_document2:
    theme: yeti
    css: fullwidth.css
    toc: true
    toc_float: true
  html_document:
    toc: true
    df_print: paged
  pdf_document:
    toc: true
    latex_engine: wkhtmltopdf
  html_notebook:
    self_contained: true
  bookdown::pdf_document2:
    toc: true
params:
  AOC_name: St. Lawrence River (Cornwall/Akwesasne)
  species: Smallmouth Bass
  contaminant: MERCURY
  aoc_shapefile: Tier3/Data/SLR_AOC
  raw_csv: "Data/Great Lakes Data to Ken 2024-12 PCB-Hg(Data).csv"
  ref_1: Lake Ontario
  ref_2: St. Lawrence River
  add_AOC: St Regis Island
  exclude_site:
  - Hamilton Harbour
  - Trent River
  - Niagara River
  target_sr: 4326
  restrict_threshold: 8
  interest_species: null
  exclude_missing: TRUE
---


```{r Setup, include = FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, messages = FALSE)


source(here::here("Scripts", "Setup.R"))
source(here::here("Tier3", "R", "assign_virtual_advisory.R"))
source(here::here("Tier3", "R", "generate_virtual_advisories.R"))

```

```{r Options, include = FALSE}
# Assigning colours to regions
region_colours <- setNames(
  c("red", "cyan", "magenta"),
  c("AOC", params$ref_1, params$ref_2)
)

# Setup Webshot 2 to browswer to allow tables to be rendered in PDF. 
# IMPORTANT: Please set this up correctly - the path to your browser application.
Sys.setenv(CHROMOTE_CHROME = "C:/Program Files/BraveSoftware/Brave-Browser/Application/brave.exe")


# Makes contaminant label title-capitalized
contaminant_label <- if (params$contaminant == "MERCURY") {
str_to_title(params$contaminant)
} else {params$contaminant}

# Get unit label
get_unit_label <- function(contaminant) {
  contaminant_upper <- toupper(contaminant)
  if (contaminant_upper == "MERCURY") {
    return("µg/g")
  } else if (contaminant_upper == "PCB") {
    return("ng/g")
  } else {
    return("units")
  }
}

unit_label <- get_unit_label(params$contaminant)


```


```{r Data-loading, include = FALSE}

raw_data = read.csv(params$raw_csv)
# Load shapefile for AOC region and reproject
aoc_shp <- st_read(params$aoc_shapefile) %>% st_transform(crs = params$target_sr)


```




```{r Dataset-Filtering, include = FALSE}




# Filter raw data by species and contaminant
filtered_data <- raw_data %>%
  filter(Specname == params$species,
         Contaminant == params$contaminant) %>%
  mutate(region = NA_character_)

has_coords <- !is.na(filtered_data$Longitude.Decimal) & !is.na(filtered_data$Latitude.Decimal)


filtered_sf <- filtered_data[has_coords, ] %>%
  st_as_sf(coords = c("Longitude.Decimal", "Latitude.Decimal"), crs = params$target_sr) %>%
  st_transform(crs = st_crs(aoc_shp)) %>%
  st_make_valid()

aoc_match <- st_intersects(filtered_sf, st_make_valid(aoc_shp), sparse = FALSE)[, 1]

# Assign AOC region to matching rows
filtered_data$region[has_coords][aoc_match] <- "AOC"


region_levels <- c("AOC", params$ref_1, if (params$ref_2 != "none") params$ref_2 else NULL)

filtered_data <- filtered_data %>%
  mutate(
    region = case_when(
      region == "AOC" ~ "AOC",  # Preserve existing spatial AOC tag
      is.na(region) & str_detect(Locname.Fishbase, fixed(params$add_AOC, ignore_case = TRUE)) ~ "AOC",
      is.na(region) & str_detect(Locname.Fishbase, fixed(params$ref_1, ignore_case = TRUE)) ~ params$ref_1,
      is.na(region) & str_detect(Locname.Fishbase, fixed(params$ref_2, ignore_case = TRUE)) ~ params$ref_2,
      TRUE ~ region
    )
  ) %>%
  filter(
    !str_detect(Locname.Fishbase, str_c(params$exclude_site, collapse = "|")),
    !is.na(region)
  ) %>%
  mutate(region = factor(region, levels = region_levels))


# Full analysis data
full_data <- if (params$exclude_missing) {
  filtered_data %>% filter(!is.na(Longitude.Decimal), !is.na(Latitude.Decimal))
} else {
  filtered_data
}

# Mapping data (recent + coordinates only)
map_data <- filtered_data %>%
  filter(Sample.Year >= 2014,
         !is.na(Longitude.Decimal),
         !is.na(Latitude.Decimal))

missing_coords <- filtered_data %>%
  filter(Sample.Year >= 2014,
         is.na(Longitude.Decimal) | is.na(Latitude.Decimal))
missing_sites <- unique(missing_coords$Locname.Fishbase) %>% sort()

recent_data <- filtered_data %>%
  filter(Sample.Year >= 2014) %>%
  { if (params$exclude_missing) filter(., !is.na(Longitude.Decimal) & !is.na(Latitude.Decimal)) else . }



filtered_data %>%
  count(region, is.na(Longitude.Decimal))  # How many sites per region w/ and w/o coords?

```

```{r legacy-filtering, eval = FALSE}
missing_coords <- filtered_data %>%
  filter(
    is.na(Longitude.Decimal) | is.na(Latitude.Decimal)
  )

missing_sites <- unique(missing_coords$Locname.Fishbase) %>% sort()


filtered_data_analysis <- if (params$exclude_missing) {
  filtered_data %>%
    filter(!is.na(Longitude.Decimal), !is.na(Latitude.Decimal))
} else {
  filtered_data
}

filtered_data_analysis <- filtered_data_analysis %>%
  mutate(region = NA_character_)

map_data <- filtered_data %>%
  filter(Sample.Year >= 2014,
         !is.na(Longitude.Decimal),
         !is.na(Latitude.Decimal))

  
  # Split into full and recent datasets
recent_data <- filtered_data %>% filter(Sample.Year >= 2014)
full_data   <- filtered_data




# Ensure geometries are valid and aligned
map_data_sf <- st_as_sf(map_data, coords = c("Longitude.Decimal", "Latitude.Decimal"), crs = params$target_sr) %>%
  st_transform(crs = st_crs(aoc_shp)) %>%
  st_make_valid()

aoc_match <- st_intersects(map_data_sf, st_make_valid(aoc_shp), sparse = FALSE)[, 1]

map_data$region <- ifelse(aoc_match, "AOC", NA)



region_levels <- c("AOC", params$ref_1, if (params$ref_2 != "none") params$ref_2 else NULL)


# Assign ref_1 and ref_2 regions using string matching
filtered_data <- filtered_data %>%
  mutate(
    region = case_when(
      region == "AOC" ~ "AOC",  # Keep existing AOC label
      is.na(region) & str_detect(Locname.Fishbase, fixed(params$add_AOC, ignore_case = TRUE)) ~ "AOC", # Add additional specified AOC sites not in geographic range of AOC
      is.na(region) & str_detect(Locname.Fishbase, fixed(params$ref_1, ignore_case = TRUE)) ~ params$ref_1,
      is.na(region) & str_detect(Locname.Fishbase, fixed(params$ref_2, ignore_case = TRUE)) ~ params$ref_2,
      TRUE ~ as.character(region)  # Preserve existing or NA
    )
  ) %>%
  filter(
    !str_detect(Locname.Fishbase, str_c(params$exclude_site, collapse = "|")),
    !is.na(region)
  ) %>%
  mutate(region = factor(region, levels = region_levels))
  



```


# Tier 3 Analysis for `r contaminant_label` in `r params$species` for the `r params$AOC_name` AOC

## Reference Sites

(ref:map) Map showing AOC and reference site locations.

```{r}
adjust_bbox_to_aspect <- function(bbox, aspect_ratio = 10/6, crs = 3857) {
  # Validate input
  if (anyNA(bbox)) {
    warning("adjust_bbox_to_aspect: bbox contains NA values.")
    return(NULL)
  }

  # Extract numeric coordinates
  xmin <- bbox["xmin"]
  xmax <- bbox["xmax"]
  ymin <- bbox["ymin"]
  ymax <- bbox["ymax"]

  width <- xmax - xmin
  height <- ymax - ymin

  if (width <= 0 || height <= 0) {
    warning("adjust_bbox_to_aspect: invalid bbox dimensions.")
    return(NULL)
  }

  current_ratio <- width / height
  center_x <- (xmin + xmax) / 2
  center_y <- (ymin + ymax) / 2

  if (current_ratio > aspect_ratio) {
    # Too wide → expand height
    new_height <- width / aspect_ratio
    new_ymin <- center_y - new_height / 2
    new_ymax <- center_y + new_height / 2
    new_bbox <- c(xmin = xmin, ymin = new_ymin, xmax = xmax, ymax = new_ymax)
  } else {
    # Too tall → expand width
    new_width <- height * aspect_ratio
    new_xmin <- center_x - new_width / 2
    new_xmax <- center_x + new_width / 2
    new_bbox <- c(xmin = new_xmin, ymin = ymin, xmax = new_xmax, ymax = ymax)
  }

  # ✅ Ensure new_bbox has correct names and is a named numeric vector
  new_bbox_named <- setNames(new_bbox, c("xmin", "ymin", "xmax", "ymax"))

  # Now construct the bbox object
  st_bbox(new_bbox_named, crs = st_crs(crs))
}



```


```{r map-creation, message = FALSE, include = FALSE}

library(maptiles)
library(ggspatial)


# Convert to sf object and project
sites_sf <- st_as_sf(map_data, coords = c("Longitude.Decimal", "Latitude.Decimal"), crs = params$target_sr)
sites_proj <- st_transform(sites_sf, 3857)  # Web Mercator

# Automatically calculate bounding box with buffer
bbox_buffered <- st_bbox(sites_proj) 
print(bbox_buffered)
bbox_adjusted <- adjust_bbox_to_aspect(bbox_buffered, aspect_ratio = 10/6, crs = 3857)
#if (!is.null(bbox_adjusted)) {
#  bbox_sfc <- st_as_sfc(bbox_adjusted) + c(-40000, -40000, 40000, 40000)
#  basemap <- get_tiles(bbox_sfc, provider = "Esri.WorldShadedRelief", crop = TRUE, zoom = 10)
#}
print(bbox_adjusted)

bbox_sfc <- st_as_sfc(bbox_adjusted +c(-40000, -40000, 40000, 40000))





# Get a clean basemap (OpenStreetMap or use Carto, Esri, etc.)
basemap <- get_tiles(bbox_sfc, provider = "Esri.WorldShadedRelief", crop = TRUE, zoom = 10)
bbox_vec <- st_bbox(basemap)

# Inset
# Filter to AOC sites
aoc_sites <- sites_proj[sites_proj$region == "AOC", ]

# Create bounding box around AOC sites with buffer
aoc_bbox <- st_bbox(aoc_sites) + c(-10000, -10000, 10000, 10000)
aoc_bbox_sfc <- st_as_sfc(aoc_bbox)

# Get basemap just for the AOC area
aoc_basemap <- get_tiles(aoc_bbox_sfc, provider = "Esri.WorldShadedRelief", crop = TRUE, zoom = 12)

# Use full set of sites for mapping
aoc_inset_sites <- sites_proj[st_within(sites_proj, aoc_bbox_sfc, sparse = FALSE), ]

```


```{r map, fig.cap='(ref:map)', message = FALSE}
aspect_ratio = 6/10

# Plot map
main_map = ggplot() +
  layer_spatial(basemap) +
  geom_sf(data = sites_proj, aes(color = region), size = 3.5, shape = 21, fill = "white", stroke = 1) +
  geom_sf(data = aoc_bbox_sfc, fill = NA, color = "red", linewidth = 0.5)+
  coord_sf(xlim = c(bbox_vec["xmin"], bbox_vec["xmax"]),
    ylim = c(bbox_vec["ymin"], bbox_vec["ymax"]),
    expand = FALSE,
    clip = "on",
    datum = NA)+
  annotation_scale(location = "bl", width_hint = 0.2) +
  #annotation_north_arrow(location = "bl", which_north = "true", style = north_arrow_fancy_orienteering) +
  labs(color = "Region") +
  theme_void(base_size = 12) +
  theme(
    legend.justification = c(1,0),
    legend.position = c(.98,.02),
    legend.box.background = element_rect(fill = "#EEFFEF", color = "black"),
    legend.box.margin = unit(c(2,2,2,2),  "pt"),
    plot.margin = unit(c(0,0,0,0), "pt")
  )+ scale_color_manual(values = region_colours)



# Create inset map
inset_map <- ggplot() +
  layer_spatial(aoc_basemap) +
  geom_sf(data = aoc_inset_sites, color = "red", size = 2.5, shape = 21, fill = "white", stroke = 0.8) +
  annotation_scale(location = "bl", width_hint = 0.3) +  # Smaller inset bar
  coord_sf(xlim = c(aoc_bbox["xmin"], aoc_bbox["xmax"]),
           ylim = c(aoc_bbox["ymin"], aoc_bbox["ymax"]),
           expand = FALSE) +
  theme_void()+ theme(aspect.ratio = NULL,panel.border = element_rect(color = "black", fill = NA))


# Combine using cowplot (inset in top-left)
library(cowplot)

combined_map <- ggdraw(main_map) +
  draw_plot(inset_map, x = -0.04, y = 0.56, width = 0.4, height = 0.35) +# adjust x/y/size as needed
  theme(plot.margin = unit(c(0,0,0,0), "pt"))

#source("Tier3/R/draw_map_with_inset.R")
#combined_map <- draw_map_with_inset(
#  main_map = main_map,
#  inset_map = inset_map,
#  pos = "tl",         # top-left
#  inset_scale = 0.35, # 35% of main map width
#  x_offset = 0.02,
#  y_offset = -0.15
#)

print(combined_map)


```


```{r Display-Reference-Sites, include = FALSE}
# Step 1: Ensure distinct region-site combinations
ref_site_listing <- recent_data %>%

  distinct(region, site = Locname.Fishbase) %>%
  arrange(region, site)

# Step 2: Collapse to one entry per region
ref_sites_md <- ref_site_listing %>%
  group_by(region) %>%
  reframe(
    markdown = paste0("#### ", unique(region), " {-}", "\n", paste0("- ", site, collapse = "\n")),
    .groups = "drop"
  ) %>%
  distinct(markdown) %>%  # <-- Remove duplicates if grouped improperly upstream
  pull(markdown) %>%
  paste(collapse = "\n\n")




```

<br />

The following reference sites were used in the Tier 3 analysis of `r contaminant_label` in `r params$species` for the `r params$AOC_names` AOC:

<br />

`r ref_sites_md`

<br />


```{r report-missing-sites-message, results = "asis"}

# Print out missing location sites

if (length(missing_sites) > 0) {
  if (params$exclude_missing) {
    cat(glue::glue("### Note\n\n {length(missing_sites)} sites with missing coordinates were excluded **from the analysis**.\n\n"))
  } else {
    cat(glue::glue("### Note\n\nRetained {length(missing_sites)} sites with missing coordinates **for analysis**, but excluded them from maps.\n\n"))
  }
  
  cat("The following sites were affected:\n\n")
  cat(paste0("- ", missing_sites, collapse = "\n"))
}


```

## Tier 3A: Virtual Advisory for `r contaminant_label` in `r params$species`

Virtual advisory values were generated using by modelling the relationship between fish length and contaminant concentration  for  `r params$species` and `r contaminant_label`. For each region, the function fits a log-log linear model (lm(log(Value) ~ log(Length)) to evaluate whether a significant relationship exists (p ≤ 0.05). If so, predicted concentrations are generated across a continuous length range, and then averaged within fixed 5 cm bins. Advisory levels are assigned using MECP-defined contaminant- and population-specific (General/Sensitive) thresholds to the mean predicted concentrations. If no significant relationship is detected, advisory levels are based on the mean concentration of all observations within each bin. 


```{r Generate virtual advisories, include = FALSE}
virtual_advisory = generate_virtual_advisories(
  df = recent_data,
  group_var = "region",
  contaminant = params$contaminant
)

```

(ref:Virtual-advisory-visualization) Virtual `r contaminant_label` advisories in `r params$species`. Predicted concentrations (curve) and observed values (points) by region using power regression model, compared to MECP advisory limits for `r params$contaminant` (dashed lines).

```{r Virtual-advisory-visualization, fig.cap='(ref:Virtual-advisory-visualization)'}
## Extract predictions and original observations
pred_df <- virtual_advisory$predictions
obs_df <- recent_data  
contaminant <- params$contaminant

y_label <- if (contaminant == "MERCURY") {
  "Mercury Concentration (µg/g)"
} else if (contaminant == "PCB") {
  "PCB Concentration (ng/g)"
} else {
  paste(contaminant_label, "(unknown units)")
}

get_thresholds <- function(contaminant = "MERCURY", population = "Sensitive") {
  if (contaminant == "MERCURY") {
    if (population == "Sensitive") {
      return(tibble(
        meals = c(32, 16, 12, 8, 4, 0),
        conc = c(0.06, 0.06, 0.12, 0.16, 0.25, 0.50)
      ))
    } else {
      return(tibble(
        meals = c(32, 16, 12, 8, 4, 2, 0),
        conc = c(0.15, 0.15, 0.3, 0.4, 0.6, 1.2, 1.8)
      ))
    }
  }
  return(NULL)
}

thresholds_df <- get_thresholds(contaminant, population = "Sensitive")


# Plot with overlaid observations and advisory thresholds
ggplot(pred_df, aes(x = Length, y = predicted, color = region)) +
  geom_point(data = obs_df, aes(x = Length, y = Value, color = region), size = 1.5, alpha = 0.7) +
  geom_line(linewidth = 1.2) +
  geom_hline(data = thresholds_df, aes(yintercept = conc), linetype = "dashed", color = "black") +
  geom_text(data = thresholds_df, aes(x = Inf, y = conc, label = paste0(meals, " meals/mo")),
            hjust = 1.1, vjust = -0.2, color = "black", size = 3, inherit.aes = FALSE) +
  scale_colour_manual(values = region_colours) +
  labs(
    x = "Length (cm)",
    y = y_label,
    color = "Region"
  ) +
  theme_minimal(base_size = 13)



```

<br />

```{r Model-summary-creation, include=FALSE}
model_summary <- virtual_advisory$model_info %>%
  mutate(
    summary_text = pmap_chr(list(lm, correlated, pval, r2), function(mod, sig, pval, r2) {
      coef_vals <- coef(mod)
      a <- signif(exp(coef_vals[1]), 3)
      b <- round(coef_vals[2], 3)
      pval_fmt <- ifelse(pval < 0.001, "< 0.001", formatC(pval, format = "e", digits = 2))
      r2_fmt <- round(r2, 3)

      sig_text <- if (sig) {
        "There is a statistically significant relationship between fish length and contaminant concentration."
      } else {
        "There is no significant relationship between fish length and contaminant concentration. Bin-level means were used instead."
      }

      model_text <- glue("The power model used was: concentration = {a} × Length<sup>{b}</sup> (R² = {r2_fmt}, p = {pval_fmt}).")

      paste0(sig_text, " ", model_text)
    })
  ) %>%
  select(region, summary_text)
```

<br />

```{r model-summary-printout, results='asis'}
# display model summaries
cat(paste0("### Model Summary by Region {-} \n\n ",
           paste0("**", model_summary$region, "**: ", model_summary$summary_text, collapse = "\n\n")))

```



 <caption>(\#tab:virtual-advisory-table) Virtual advisory levels for `r contaminant_label` in `r params$species`  across regions. Values represent meals/month recommendations.</caption>

```{r virtual-advisory-table, results='asis', warning = TRUE}

predictions_binned <- virtual_advisory$predictions %>%
  mutate(
    length_bin = cut(
      Length,
      breaks = c(seq(15, 75, by = 5), Inf),
      labels = c(paste(seq(15, 70, by = 5), seq(20, 75, by = 5), sep = "-"), "75+"),
      right = FALSE
    )
  ) %>%
  filter(!is.na(length_bin)) %>%
  group_by(region, length_bin) %>%
  summarise(
    mean_conc = mean(predicted, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  mutate(
    meals_general = assign_advisory(mean_conc, "General", toupper(params$contaminant)),
    meals_sensitive = assign_advisory(mean_conc, "Sensitive", toupper(params$contaminant))
  )



color_advisory_cell <- function(value, threshold = params$restrict_threshold) {
  if (is.na(value)) return("")
  if (value < threshold) {
    return(list(background = "red", color = "white", fontWeight = "bold"))
  } else {
    return(list(background = "forestgreen", color = "black", fontWeight = "bold"))
  }
}



virtual_long <- predictions_binned %>%
  pivot_longer(cols = c(meals_general, meals_sensitive),
               names_to = "Population",
               names_prefix = "meals_",
               values_to = "Advisory") %>%
  mutate(
    Population = case_when(
      Population == "general" ~ "General",
      Population == "sensitive" ~ "Sensitive"
    )
  )


virtual_wide <- virtual_long %>%
  select(Region = region, Population, length_bin, Advisory) %>%
  pivot_wider(names_from = length_bin, values_from = Advisory)


virtual_adv_table <- reactable(
  virtual_wide,
  bordered = TRUE,
  highlight = TRUE,
  striped = TRUE,
  defaultColDef = colDef(
    style = function(value) {
      if (is.numeric(value) && !is.na(value)) {
        color_advisory_cell(value)
      } else {
        NULL
      }
    },
    align = "center"
  ),
  columns = list(
    Region = colDef(name = "Region", sticky = "left", style = list(fontWeight = "bold")),
    Population = colDef(name = "Population", sticky = "left", style = list(fontWeight = "bold"))
  )
)

#htmlwidgets::saveWidget(virtual_adv_table, "virtual_adv_table.html", selfcontained = TRUE)
# Screenshot the widget for PDF
#webshot2::webshot(
#  "virtual_adv_table.html",
#  file = "virtual_adv_table.png",
#  selector = "body",
#  zoom = 2,
#  vheight = 1200,  # increase this to fit longer tables
#  cliprect = "viewport"
#)

virtual_adv_table

```


<br />


## Tier 3B: Comparison of contaminant concentrations in `r params$species` in AOC to reference

To assess differences in fish contamination between the AOC and reference sites, contaminant–length relationships were modeled using Generalized Additive Models (GAMs), allowing for non-linear associations. For each priority contaminant and indicator species, we first tested whether the shape of the concentration–length relationship differed significantly between AOC and reference fish. Where models supported statistically similar trends (non-significant interaction), we compared size-adjusted contaminant levels using estimated marginal means. Where shapes differed significantly, contaminant concentration values were compared by GAM-predicted values at standardized lengths (e.g., 45cm) commonly consumed for each fish species. A threshold of α = 0.05 was used to evaluate statistical significance throughout.

(ref:Tier-3B-Figure) `r contaminant_label` concentration in `r params$species` by length in the AOC compared to reference sites.

```{r Tier-3B-Figure, echo = FALSE, message = FALSE, fig.cap='(ref:Tier-3B-Figure)'}
# Hg concentration of fish by size
ggplot(recent_data, aes(x = Length, y = Value, color = region)) +
  geom_point(alpha = 0.9) +
  geom_smooth(method = "gam", se = TRUE, linewidth = 1.2) +
  labs(
    title = "Contaminant Concentration vs. Fish Length",
    x = "Length (cm)",
    y = y_label
  ) +
  theme_minimal(base_size = 14) +
  scale_colour_manual(values = region_colours)



```




```{r Tier 3B Reference Comparison}


# Define standard comparison length
standard_length <- 30

# Fit GAM with region-specific smooths
gam_model <- gam(Value ~ s(Length, by = region) + region,
                 data = recent_data,
                 method = "REML")

# summary for report
gam_summary <- summary(gam_model)

# Test for interaction between Length and region (i.e., different smooths)
# Compare with reduced model with shared smooth
gam_null <- gam(Value ~ s(Length) + region,
                data = recent_data,
                method = "REML")

interaction_test <- anova(gam_null, gam_model, test = "Chisq")
interaction_p <- interaction_test$`Pr(>Chi)`[2]

# Predict contaminant concentration at standard length
# Create new data frame for prediction
new_data <- data.frame(
  Length = rep(standard_length, 3),
  region = c("AOC", params$ref_1, params$ref_2)
)

# Predict with standard errors
predicted <- predict(gam_model, new_data, se.fit = TRUE)
pred_df <- new_data %>%
  mutate(
    fit = predicted$fit,
    se = predicted$se.fit,
    lower = fit - 1.96 * se,
    upper = fit + 1.96 * se
  )

# Post-hoc comparison at standard length using emmeans
em <- emmeans(gam_model, ~ region | Length,
              at = list(Length = standard_length),
              type = "response")

# Pairwise comparisons between regions (Tukey adjustment optional)
em_contrasts <- contrast(em, method = "pairwise", adjust = "none") %>% summary()


```

```{r}
# Make region label nicer
ref_label <- ifelse(params$ref_2 != "none", paste(params$ref_1, "and", params$ref_2), params$ref_1)

# Extract p-value and format it
interaction_p_val <- signif(interaction_p, 3)

if (interaction_p < 0.05) {
  # Filter predictions
  pred_aoc <- pred_df %>% filter(region == "AOC")
  pred_refs <- pred_df %>% filter(region != "AOC")

  # Build reference comparison text
  ref_texts <- pred_refs %>%
    mutate(
      out = glue::glue("{region}: {round(fit, 2)} {unit_label} (95% CI: {round(lower, 2)}–{round(upper, 2)})")
    ) %>%
    pull(out) %>%
    paste(collapse = "; ")

  summary_text <- glue::glue(
    "Generalized Additive Models (GAMs) were used to model the relationship between fish length and {tolower(params$contaminant)} concentrations for {params$species} in the AOC and reference sites ({paste(region_levels[-1], collapse = ', ')}). ",
    "The length-contaminant relationship differed significantly between sites (interaction test p = {interaction_p_val}). ",
    "Model-predicted concentrations were compared at a standardized fish length of {standard_length} cm. ",
    "At this length, the predicted {tolower(params$contaminant)} concentration was {round(pred_aoc$fit, 2)} {unit_label} (95% CI: {round(pred_aoc$lower, 2)}–{round(pred_aoc$upper, 2)}) in the AOC. ",
    "Predicted concentrations in reference site(s) were: {ref_texts}."
  )

} else {
  # Get emmeans for each region
  em_summary <- summary(emmeans(gam_model, ~ region))
  em_aoc <- em_summary %>% filter(region == "AOC")
  em_refs <- em_summary %>% filter(region != "AOC")

  # Build reference comparison text
  ref_texts <- em_refs %>%
    mutate(
      out = glue::glue("{region}: {round(emmean, 2)} {unit_label} (95% CI: {round(lower.CL, 2)}–{round(upper.CL, 2)})")
    ) %>%
    pull(out) %>%
    paste(collapse = "; ")

  summary_text <- glue::glue(
    "Generalized Additive Models (GAMs) were used to model the relationship between fish length and {tolower(params$contaminant)} concentrations for {params$species} in the AOC and reference regions ({paste(region_levels[-1], collapse = ', ')}). ",
    "The length-contaminant relationship did not differ significantly between sites (interaction test p = {interaction_p_val}). ",
    "Size-adjusted marginal means were compared using the GAM model. ",
    "The estimated marginal mean {tolower(params$contaminant)} concentration in the AOC was {round(em_aoc$emmean, 2)} {unit_label} (95% CI: {round(em_aoc$lower.CL, 2)}–{round(em_aoc$upper.CL, 2)}). ",
    "Estimated values for reference site(s): {ref_texts}."
  )
}



```

<br />

`r summary_text`


```{r tier 3b summary table, include=FALSE}
library(knitr)

# Table of predicted or marginal means
summary_table <- if (interaction_p < 0.05) {
  pred_df %>%
    mutate(
      `Predicted Value ({unit_label})` = round(fit, 2),
      `95% CI` = paste0(round(lower, 2), "–", round(upper, 2))
    ) %>%
    select(Region = region, `Predicted Value ({unit_label})`, `95% CI`)
} else {
  em_summary %>%
    mutate(
      `Estimated Mean ({unit_label})` = round(emmean, 2),
      `95% CI` = paste0(round(lower.CL, 2), "–", round(upper.CL, 2))
    ) %>%
    select(Region = region, `Estimated Mean ({unit_label})`, `95% CI`)
}

kable(summary_table, caption = "Summary of predicted or adjusted contaminant concentrations by region.")

```


## Tier 3C: Temporal Trends

Temporal trends in `r params$species` for `r contaminant_label` were analyzed across all years using Generalized Additive Models (GAMs). The analysis was limited to the three most sampled length bins for the species, and results are presented separately by region. These plots help assess whether contaminant concentrations are declining, stable, or increasing over time within each region.


```{r temp-trends}

# Create length bins
temporal_df <- full_data %>%
  mutate(length_bin = cut(
    Length,
    breaks = c(seq(25, 75, by = 5), Inf),
    labels = c(paste(seq(25, 70, by = 5), seq(29, 74, by = 5), sep = "-"), "75+"),
    right = FALSE
  ))


# Identify top 3 most sampled bins
top_bins <- temporal_df %>%
  count(length_bin, sort = TRUE) %>%
  slice_head(n = 3) %>%
  pull(length_bin)

# Filter to those bins only
temporal_top_bins <- temporal_df %>%
  filter(length_bin %in% top_bins)
```

(ref:temp-trends-fig) Temporal `r contaminant_label` concentration (`r unit_label`) trends for `r params$species`, grouped by region and top 3 sampled size bins. The top 3 most sampled size bins for `r params$species` in the regions of interest are `r glue::glue_collapse(top_bins, sep = ", ", last = " and ")` cm, in that order. Lines represent GAM-smooth temporal trends.

```{r temp-trends-fig, fig.cap='(ref:temp-trends-fig)', fig.width=8,fig.height=14, message=FALSE}
# Function to create a trend plot per region
plot_region_trends <- function(data, region_label) {
  ggplot(data, aes(x = Sample.Year, y = Value, color = length_bin, shape = length_bin)) +
    geom_point(alpha = 0.7, size = 2) +
    geom_smooth(method = "gam", se = TRUE, linewidth = 1) +
    labs(
     # title = paste("Temporal", contaminant_label, "Trends in", params$species, "-", region_label),
     # subtitle = paste("Top 3 length bins (cm):", paste(top_bins, collapse = ", ")),
      x = "Sample Year",
      y = paste(contaminant_label, unit_label),
      color = "Length Bin (cm)",
      shape = "Length Bin (cm)"
    ) +
    scale_color_manual(values = c("red", "magenta", "cyan")) +
    theme_classic(base_size = 13)
}

# Generate plots for each region
aoc_plot <- plot_region_trends(filter(temporal_top_bins, region == "AOC"), "AOC") + ggtitle(params$AOC_name)
ref1_plot <- plot_region_trends(filter(temporal_top_bins, region == params$ref_1), params$ref_1)+ ggtitle(params$ref_1)
ref2_plot  <- plot_region_trends(filter(temporal_top_bins, region == params$ref_2), params$ref_2)+ ggtitle(params$ref_2)



library(ggpubr)
ggarrange(aoc_plot, ref1_plot, ref2_plot,
          ncol = 1,
          nrow = 3,
          align = "h",
          common.legend = TRUE,
          legend = "bottom"
          )
```






```{r, eval = FALSE}
all_ontario_sites = c(    -'44107646'
    -"43237942"
    -"43187947"
    -"44127632"
    -"44147628"
    -"43137937"
    -"44117644"
    -"43267940"
    -"43157915"
    -"43567817"
    -"43037913"
    -"43567720"
    -"44067627"
    -"44077656"
    -"44107644"
    -"43247953"
    -"43127917"
    -"43177906"
    -"43357917"
    -"43347930"
    -"43407919"
    -"43377929"
    -"43367928"
    -"43377923"
    -"43397919"
    -"43357943"
    -"43497859"
    -"43577809"
    -"43507840"
    -"43487910"
    -"43497905"
    -"43527856"
    -"43577818"
    -"43567717"
    -"44017743"
    -"44067734"
    -"44097717"
    -"44067735"
    -"44057733"
    -"44097721"
    -"44087724"
    )