---
title: "Tier 3 Report - SLR"
author: "Aaron Dain"
always_allow_html: true
output:
  bookdown::html_document2:
    theme: yeti
    css: fullwidth.css
    toc: true
    toc_float: true
  html_document:
    toc: true
    df_print: paged
  pdf_document:
    toc: true
    latex_engine: wkhtmltopdf
  html_notebook:
    self_contained: true
  bookdown::pdf_document2:
    toc: true
params:
  AOC_name: St. Lawrence River (Cornwall/Akwesasne)
  species: Yellow Perch
  contaminant: MERCURY
  aoc_shapefile: Tier3/Data/SLR_AOC
  raw_csv: "Data/Great Lakes Data to Ken 2024-12 PCB-Hg(Data).csv"
  ref_1: Lake Ontario
  ref_2: St. Lawrence River
  combine_ref: TRUE
  add_AOC: St Regis Island
  exclude_site:
  - Hamilton Harbour
  - Trent River
  - Niagara River
  target_sr: 4326
  restrict_threshold: 8
  exclude_missing: TRUE
  standard_length: 45
---

```{r Setup, include = FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, messages = FALSE)


source(here::here("Scripts", "Setup.R"))

raw_data = read.csv(params$raw_csv)

tool_files_t3 = list.files(path = "Tier3/R", pattern = "\\.R$", full.names = TRUE)
invisible(lapply(tool_files_t3, source))



# Dynamically assign ref_1 and ref_2 based on whether combining reference regions
if (params$combine_ref) {
  ref_1 <- "Reference"
  ref_2 <- "none"
  region_levels <- c("AOC", ref_1)
  ref_label <- "Reference"
  ref_regions_to_match <- c(params$ref_1, params$ref_2)
} else {
  ref_1 <- params$ref_1
  ref_2 <- params$ref_2
  region_levels <- c("AOC", ref_1, if (ref_2 != "none") ref_2 else NULL)
  ref_label <- if (ref_2 != "none") paste(ref_1, "and", ref_2) else ref_1
  ref_regions_to_match <- c(ref_1, if (ref_2 != "none") ref_2 else NULL)
}


# Check if AOC shapefile exists
use_aoc_shapefile <- !is.null(params$aoc_shapefile) && file.exists(params$aoc_shapefile) 

```

```{r Options, include = FALSE}

# Assigning colours to regions
region_colours <- setNames(
  c("red", "cyan", "magenta")[seq_along(region_levels)],
  region_levels
)



# Setup Webshot 2 to browswer to allow tables to be rendered in PDF. 
# IMPORTANT: Please set this up correctly - the path to your browser application.
Sys.setenv(CHROMOTE_CHROME = "C:/Program Files/BraveSoftware/Brave-Browser/Application/brave.exe")


# Makes contaminant label title-capitalized
contaminant_label <- if (params$contaminant == "MERCURY") {
str_to_title(params$contaminant)
} else {params$contaminant}

# Get unit label
get_unit_label <- function(contaminant) {
  contaminant_upper <- toupper(contaminant)
  if (contaminant_upper == "MERCURY") {
    return("µg/g")
  } else if (contaminant_upper == "PCBs") {
    return("ng/g")
  } else {
    return("units")
  }
}

unit_label <- get_unit_label(params$contaminant)



```

```{r Dataset-Filtering, include = FALSE}


# Filter raw data by species and contaminant
filtered_data <- raw_data %>%
  filter(Specname == params$species,
         Contaminant == params$contaminant) %>%
  mutate(region = NA_character_)

has_coords <- !is.na(filtered_data$Longitude.Decimal) & !is.na(filtered_data$Latitude.Decimal)

# Load shapefile for AOC region and reproject
if (use_aoc_shapefile) {
  aoc_shp <- st_read(params$aoc_shapefile) %>% st_transform(crs = params$target_sr)

  filtered_sf <- filtered_data[has_coords, ] %>%
    st_as_sf(coords = c("Longitude.Decimal", "Latitude.Decimal"), crs = params$target_sr) %>%
    st_transform(crs = st_crs(aoc_shp)) %>%
    st_make_valid()

  aoc_match <- st_intersects(filtered_sf, st_make_valid(aoc_shp), sparse = FALSE)[, 1]
  filtered_data$region[has_coords][aoc_match] <- "AOC"
}



filtered_data <- filtered_data %>%
  mutate(
    region = case_when(
      region == "AOC" ~ "AOC",
      is.na(region) & str_detect(Locname.Fishbase, regex(str_c(params$add_AOC, collapse = "|"), ignore_case = TRUE)) ~ "AOC",

      params$combine_ref & is.na(region) & (
        str_detect(Locname.Fishbase, fixed(params$ref_1, ignore_case = TRUE)) |
        str_detect(Locname.Fishbase, fixed(params$ref_2, ignore_case = TRUE))
      ) ~ "Reference",

      !params$combine_ref & is.na(region) & str_detect(Locname.Fishbase, fixed(params$ref_1, ignore_case = TRUE)) ~ params$ref_1,

      !params$combine_ref & is.na(region) & 
        !is.null(params$ref_2) & 
        params$ref_2 != "none" &
        str_detect(Locname.Fishbase, fixed(params$ref_2, ignore_case = TRUE)) ~ params$ref_2,

      TRUE ~ region
    )
  ) %>%
  filter(
    !str_detect(Locname.Fishbase, str_c(params$exclude_site, collapse = "|")),
    !is.na(region)
  ) %>%
  mutate(region = factor(region, levels = region_levels))


# Full analysis data
full_data <- if (params$exclude_missing) {
  filtered_data %>% filter(!is.na(Longitude.Decimal), !is.na(Latitude.Decimal))
} else {
  filtered_data
}

# Mapping data (recent + coordinates only)
map_data <- filtered_data %>%
  filter(Sample.Year >= 2014,
         !is.na(Longitude.Decimal),
         !is.na(Latitude.Decimal))

missing_coords <- filtered_data %>%
  filter(Sample.Year >= 2014,
         is.na(Longitude.Decimal) | is.na(Latitude.Decimal))
missing_sites <- unique(missing_coords$Locname.Fishbase) %>% sort()

recent_data <- filtered_data %>%
  filter(Sample.Year >= 2014) %>%
  { if (params$exclude_missing) filter(., !is.na(Longitude.Decimal) & !is.na(Latitude.Decimal)) else . }



filtered_data %>%
  count(region, is.na(Longitude.Decimal))  # How many sites per region w/ and w/o coords?

```

# Tier 3 Analysis for `r contaminant_label` in `r params$species` for the `r params$AOC_name` AOC

## Reference Sites

(ref:map) Map showing AOC and reference site locations.

```{r map-setup}
adjust_bbox_to_aspect <- function(bbox, aspect_ratio = 10/6, crs = 3857) {
  # Validate input
  if (anyNA(bbox)) {
    warning("adjust_bbox_to_aspect: bbox contains NA values.")
    return(NULL)
  }

  # Extract numeric coordinates
  xmin <- bbox["xmin"]
  xmax <- bbox["xmax"]
  ymin <- bbox["ymin"]
  ymax <- bbox["ymax"]

  width <- xmax - xmin
  height <- ymax - ymin

  if (width <= 0 || height <= 0) {
    warning("adjust_bbox_to_aspect: invalid bbox dimensions.")
    return(NULL)
  }

  current_ratio <- width / height
  center_x <- (xmin + xmax) / 2
  center_y <- (ymin + ymax) / 2

  if (current_ratio > aspect_ratio) {
    # Too wide → expand height
    new_height <- width / aspect_ratio
    new_ymin <- center_y - new_height / 2
    new_ymax <- center_y + new_height / 2
    new_bbox <- c(xmin = xmin, ymin = new_ymin, xmax = xmax, ymax = new_ymax)
  } else {
    # Too tall → expand width
    new_width <- height * aspect_ratio
    new_xmin <- center_x - new_width / 2
    new_xmax <- center_x + new_width / 2
    new_bbox <- c(xmin = new_xmin, ymin = ymin, xmax = new_xmax, ymax = ymax)
  }

  # ✅ Ensure new_bbox has correct names and is a named numeric vector
  new_bbox_named <- setNames(new_bbox, c("xmin", "ymin", "xmax", "ymax"))

  # Now construct the bbox object
  st_bbox(new_bbox_named, crs = st_crs(crs))
}



```

```{r map-creation, message = FALSE, include = FALSE}

library(maptiles)
library(ggspatial)

# Convert to sf object and project
sites_sf <- st_as_sf(map_data, coords = c("Longitude.Decimal", "Latitude.Decimal"), crs = params$target_sr)
sites_proj <- st_transform(sites_sf, 3857)  # Web Mercator

# Main map bounding box
bbox_buffered <- st_bbox(sites_proj)
bbox_adjusted <- adjust_bbox_to_aspect(bbox_buffered, aspect_ratio = 10/6, crs = 3857)
bbox_sfc <- st_as_sfc(bbox_adjusted + c(-40000, -40000, 40000, 40000))

# Get basemap
basemap <- get_tiles(bbox_sfc, provider = "Esri.WorldShadedRelief", crop = TRUE, zoom = 10)
bbox_vec <- st_bbox(basemap)

# -------- AOC inset logic (only if shapefile is being used) --------
aoc_sites <- sites_proj[sites_proj$region == "AOC", ]
has_inset <- nrow(aoc_sites) > 0

if (has_inset) {
  aoc_bbox <- st_bbox(aoc_sites) + c(-10000, -10000, 10000, 10000)
  aoc_bbox_sfc <- st_as_sfc(aoc_bbox)
  aoc_basemap <- get_tiles(aoc_bbox_sfc, provider = "Esri.WorldShadedRelief", crop = TRUE, zoom = 12)

  aoc_inset_sites <- sites_proj[st_within(sites_proj, aoc_bbox_sfc, sparse = FALSE), ]
}


```

```{r map, fig.cap='(ref:map)', message = FALSE}
aspect_ratio = 6/10

# Plot map
main_map = ggplot() +
  layer_spatial(basemap) +
  geom_sf(data = sites_proj, aes(color = region), size = 3.5, shape = 21, fill = "white", stroke = 1) +
  geom_sf(data = aoc_bbox_sfc, fill = NA, color = "red", linewidth = 0.5)+
  coord_sf(xlim = c(bbox_vec["xmin"], bbox_vec["xmax"]),
    ylim = c(bbox_vec["ymin"], bbox_vec["ymax"]),
    expand = FALSE,
    clip = "on",
    datum = NA)+
  annotation_scale(location = "bl", width_hint = 0.2) +
  #annotation_north_arrow(location = "bl", which_north = "true", style = north_arrow_fancy_orienteering) +
  labs(color = "Region") +
  theme_void(base_size = 12) +
  theme(
    legend.justification = c(1,0),
    legend.position = c(.98,.02),
    legend.box.background = element_rect(fill = "#EEFFEF", color = "black"),
    legend.box.margin = unit(c(2,2,2,2),  "pt"),
    plot.margin = unit(c(0,0,0,0), "pt")
  )+ scale_color_manual(values = region_colours)



# Create inset map
inset_map <- ggplot() +
  layer_spatial(aoc_basemap) +
  geom_sf(data = aoc_inset_sites, aes(color = region), size = 2.5, shape = 21, fill = "white", stroke = 0.8) +
  annotation_scale(location = "bl", width_hint = 0.3) +
  coord_sf(xlim = c(aoc_bbox["xmin"], aoc_bbox["xmax"]),
           ylim = c(aoc_bbox["ymin"], aoc_bbox["ymax"]),
           expand = FALSE) +
  scale_colour_manual(values = region_colours) +  
  theme_void() +
  theme(
    aspect.ratio = NULL,
    panel.border = element_rect(color = "black", fill = NA),
    legend.position = "none"  # Suppress inset legend
  )



# Combine using cowplot (inset in top-left)
library(cowplot)

combined_map <- ggdraw(main_map) +
  draw_plot(inset_map, x = -0.04, y = 0.6, width = 0.4, height = 0.35) +# adjust x/y/size as needed
  theme(plot.margin = unit(c(0,0,0,0), "pt"))

#source("Tier3/R/draw_map_with_inset.R")
#combined_map <- draw_map_with_inset(
#  main_map = main_map,
#  inset_map = inset_map,
#  pos = "tl",         # top-left
#  inset_scale = 0.35, # 35% of main map width
#  x_offset = 0.02,
#  y_offset = -0.15
#)

print(combined_map)


```

```{r Display-Reference-Sites, include = FALSE}
# Step 1: Ensure distinct region-site combinations
ref_site_listing <- recent_data %>%

  distinct(region, site = Locname.Fishbase) %>%
  arrange(region, site)

# Step 2: Collapse to one entry per region
ref_sites_md <- ref_site_listing %>%
  group_by(region) %>%
  reframe(
    markdown = paste0("#### ", unique(region), " {-}", "\n", paste0("- ", site, collapse = "\n")),
    .groups = "drop"
  ) %>%
  distinct(markdown) %>%  # <-- Remove duplicates if grouped improperly upstream
  pull(markdown) %>%
  paste(collapse = "\n\n")




```

<br />

The following reference sites were used in the Tier 3 analysis of `r contaminant_label` in `r params$species` for the `r params$AOC_names` AOC:

<br />

`r ref_sites_md`

<br />

```{r report-missing-sites-message, results = "asis"}

# Print out missing location sites

if (length(missing_sites) > 0) {
  if (params$exclude_missing) {
    cat(glue::glue("### Note\n\n {length(missing_sites)} sites with missing coordinates were excluded **from the analysis**.\n\n"))
  } else {
    cat(glue::glue("### Note\n\nRetained {length(missing_sites)} sites with missing coordinates **for analysis**, but excluded them from maps.\n\n"))
  }
  
  cat("The following sites were affected:\n\n")
  cat(paste0("- ", missing_sites, collapse = "\n"))
}


```

## Tier 3A: Virtual Advisory for `r contaminant_label` in `r params$species`

Virtual advisory values were generated using by modelling the relationship between fish length and contaminant concentration for `r params$species` and `r contaminant_label`. For each region, the function fits a log-log linear model (lm(log(Value) \~ log(Length)) to evaluate whether a significant relationship exists (p ≤ 0.05). If so, predicted concentrations are generated across a continuous length range, and then averaged within fixed 5 cm bins. Advisory levels are assigned using MECP-defined contaminant- and population-specific (General/Sensitive) thresholds to the mean predicted concentrations. If no significant relationship is detected, advisory levels are based on the mean concentration of all observations within each bin.

```{r Generate virtual advisories, include = FALSE}
virtual_advisory = generate_virtual_advisories(
  df = recent_data,
  group_var = "region",
  contaminant = params$contaminant
)

```

(ref:Virtual-advisory-visualization) Virtual `r contaminant_label` advisories in `r params$species`. Predicted concentrations (curve) and observed values (points) by region using power regression model, compared to MECP advisory limits for `r params$contaminant` (dashed lines).

```{r Virtual-advisory-visualization, fig.cap='(ref:Virtual-advisory-visualization)'}
# Extract predictions and original observations
pred_df <- virtual_advisory$predictions
obs_df <- recent_data  
contaminant <- params$contaminant

y_label <- if (contaminant == "MERCURY") {
  "Mercury Concentration (µg/g)"
} else if (contaminant == "PCBs") {
  "PCBs Concentration (ng/g)"
} else {
  paste(contaminant_label, "(unknown units)")
}

get_thresholds <- function(contaminant = "MERCURY", population = "Sensitive") {
  if (contaminant == "MERCURY") {
    if (population == "Sensitive") {
      return(tibble(
        meals = c(16, 12, 8, 4, 0),
        conc = c(0.06, 0.12, 0.16, 0.25, 0.50)
      ))
    } else {
      return(tibble(
        meals = c(32, 16, 12, 8, 4, 2, 0),
        conc = c(0.15, 0.15, 0.3, 0.4, 0.6, 1.2, 1.8)
      ))
    }
  } else if (contaminant == "PCBs") {
    return(tibble(
      meals = c(16, 12, 8, 4, 2, 1, 0),
      conc = c(26, 53, 70, 105, 211, 422, 844)
    ))
  
} else {
    
  return(NULL)
}
}

thresholds_df <- get_thresholds(contaminant, population = "Sensitive")


# Plot with overlaid observations and advisory thresholds
ggplot(pred_df, aes(x = Length, y = predicted, color = region)) +
  geom_point(data = obs_df, aes(x = Length, y = Value, color = region), size = 1.5, alpha = 0.7) +
  geom_line(linewidth = 1.2) +
  geom_hline(data = thresholds_df, aes(yintercept = conc), linetype = "dashed", color = "black") +
  geom_text(data = thresholds_df, aes(x = Inf, y = conc, label = paste0(meals, " meals/mo")),
            hjust = 1.1, vjust = -0.2, color = "black", size = 3, inherit.aes = FALSE) +
  scale_colour_manual(values = region_colours) +
  labs(
    x = "Length (cm)",
    y = y_label,
    color = "Region"
  ) +
  theme_minimal(base_size = 13)



```

<br />

```{r Model-summary-creation, include=FALSE}
model_summary <- virtual_advisory$model_info %>%
  mutate(
    summary_text = pmap_chr(list(lm, correlated, pval, r2), function(mod, sig, pval, r2) {
      coef_vals <- coef(mod)
      a <- signif(exp(coef_vals[1]), 3)
      b <- round(coef_vals[2], 3)
      pval_fmt <- ifelse(pval < 0.001, "< 0.001", formatC(pval, format = "e", digits = 2))
      r2_fmt <- round(r2, 3)

      sig_text <- if (sig) {
        "There is a statistically significant relationship between fish length and contaminant concentration."
      } else {
        "There is no significant relationship between fish length and contaminant concentration. Bin-level means were used instead."
      }

      model_text <- glue("The power model used was: concentration = {a} × Length<sup>{b}</sup> (R² = {r2_fmt}, p = {pval_fmt}).")

      paste0(sig_text, " ", model_text)
    })
  ) %>%
  select(region, summary_text)
```

<br />

```{r model-summary-printout, results='asis'}
# display model summaries
cat(paste0("### Model Summary by Region {-} \n\n ",
           paste0("**", model_summary$region, "**: ", model_summary$summary_text, collapse = "\n\n")))

```

<caption>(#tab:virtual-advisory-table) Virtual advisory levels for `r contaminant_label` in `r params$species` across regions. Values represent meals/month recommendations. Red cells indicate restrictive virtual advisory falling below the desired consumption rate of `r params$restrict_threshold` meals/month; green cells indicate unrestrictive virtual advisory equal to or above the desired consumption rate.</caption>

```{r virtual-advisory-table, results='asis', warning = TRUE}

predictions_binned <- virtual_advisory$predictions %>%
  mutate(
    length_bin = cut(
      Length,
      breaks = c(seq(15, 75, by = 5), Inf),
      labels = c(paste(seq(15, 70, by = 5), seq(20, 75, by = 5), sep = "-"), "75+"),
      right = FALSE
    )
  ) %>%
  filter(!is.na(length_bin)) %>%
  group_by(region, length_bin) %>%
  summarise(
    mean_conc = mean(predicted, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  mutate(
  meals_general   = assign_advisory(mean_conc, "General", params$contaminant),
  meals_sensitive = assign_advisory(mean_conc, "Sensitive", params$contaminant)
)





color_advisory_cell <- function(value, threshold = params$restrict_threshold) {
  if (is.na(value)) return("")
  if (value < threshold) {
    return(list(background = "red", color = "white", fontWeight = "bold"))
  } else {
    return(list(background = "forestgreen", color = "black", fontWeight = "bold"))
  }
}



virtual_long <- predictions_binned %>%
  pivot_longer(cols = c(meals_general, meals_sensitive),
               names_to = "Population",
               names_prefix = "meals_",
               values_to = "Advisory") %>%
  mutate(
    Population = case_when(
      Population == "general" ~ "General",
      Population == "sensitive" ~ "Sensitive"
    )
  )%>%
  filter(
    toupper(params$contaminant) == "MERCURY" | Population == "General"
  )


virtual_wide <- virtual_long %>%
  select(Region = region, Population, length_bin, Advisory) %>%
  pivot_wider(names_from = length_bin, values_from = Advisory)


virtual_adv_table <- reactable(
  virtual_wide,
  bordered = TRUE,
  highlight = TRUE,
  striped = TRUE,
  defaultColDef = colDef(
    style = function(value) {
      if (is.numeric(value) && !is.na(value)) {
        color_advisory_cell(value)
      } else {
        NULL
      }
    },
    align = "center"
  ),
  columns = list(
    Region = colDef(name = "Region", sticky = "left", style = list(fontWeight = "bold")),
    Population = colDef(name = "Population", sticky = "left", style = list(fontWeight = "bold"))
  )
)


virtual_adv_table
```

```{r tier-3a-decision}
# Decision making


# Step 3: Is AOC more restrictive than reference? Split AOC and reference tables
aoc_advisories <- virtual_long %>%
  filter(region == "AOC") %>%
  select(length_bin, Population, Advisory) %>%
  rename(AOC_Advisory = Advisory)

reference_advisories <- virtual_long %>%
  filter(region != "AOC") %>%
  group_by(length_bin, Population) %>%
  summarise(Ref_Advisory = median(Advisory, na.rm = TRUE), .groups = "drop")  # Median across references - important if >1 reference group


# Join and evaluate
tier3a_joined <- aoc_advisories %>%
  left_join(reference_advisories, by = c("length_bin", "Population")) %>%
  mutate(
    restrict_pass = AOC_Advisory >= params$restrict_threshold,
    restrict_fail = !restrict_pass,
    more_restrictive_than_ref = AOC_Advisory < Ref_Advisory,
    final_decision = case_when(
      restrict_pass ~ "Pass",
      restrict_fail & !more_restrictive_than_ref ~ "Pass",
      restrict_fail & more_restrictive_than_ref ~ "Fail",
      TRUE ~ "Unknown"
    )
  )


tier3a_summary <- tier3a_joined %>%
  group_by(Population) %>%
  summarise(
    n_bins = n(),
    n_below_thresh = sum(restrict_fail),
    n_below_ref = sum(restrict_fail & more_restrictive_than_ref),
    n_pass = n_bins - n_below_ref,
    pct_pass = round(100 * n_pass / n_bins, 1),
    .groups = "drop"
  ) %>%
  mutate(
    decision = if_else(n_below_ref == 0, "Pass", "Fail")
  )



tier3a_text <- tier3a_summary %>%
  mutate(
    summary = glue::glue(
      "**{Population}** population: Of {n_bins} size bins, {n_below_thresh} were below the desired consumption rate of {params$restrict_threshold} meals/month. ",
      "Of these, {n_below_ref} were more restrictive than reference."
    )
  ) %>%
  pull(summary) %>%
  glue::glue_collapse(sep = " ")

tier3a_overall <- if (any(tier3a_summary$decision == "Fail")) "Fail" else "Pass"


tier3a_summary_tibble <- tibble(
  Tier = "Tier 3A",
  Description = "Virtual Advisory",
  Metric = glue_collapse(glue("{tier3a_summary$Population}: {tier3a_summary$n_below_ref}/{tier3a_summary$n_bins} size bins restrictive or more restrictive than reference"), sep = "; "),
  Outcome = tier3a_overall
)
```

```{r tier-3a-decision-table, results='asis'}
cat(tier3a_text)

```

<br />

**Tier 3A Overall Decision:** `r tier3a_overall`

<br />

## Tier 3B: Comparison of contaminant concentrations in `r params$species` in AOC to reference

To assess differences in fish contamination between the AOC and reference sites, contaminant–length relationships were modeled using Generalized Additive Models (GAMs), allowing for non-linear associations. For each priority contaminant and indicator species, we first tested whether the shape of the concentration–length relationship differed significantly between AOC and reference fish. Where models supported statistically similar trends (non-significant interaction), we compared size-adjusted contaminant levels using estimated marginal means. Where shapes differed significantly, contaminant concentration values were compared by GAM-predicted values at standardized lengths (e.g., 45cm) commonly consumed for each fish species. A threshold of α = 0.05 was used to evaluate statistical significance throughout.

(ref:Tier-3B-Figure) `r contaminant_label` concentration in `r params$species` by length in the AOC compared to reference sites.

```{r Tier-3B-Figure, echo = FALSE, message = FALSE, fig.cap='(ref:Tier-3B-Figure)'}
# Hg concentration of fish by size
ggplot(recent_data, aes(x = Length, y = Value, color = region)) +
  geom_point(alpha = 0.9) +
  geom_smooth(method = "gam", se = TRUE, linewidth = 1.2) +
  labs(
    title = "Contaminant Concentration vs. Fish Length",
    x = "Length (cm)",
    y = y_label
  ) +
  theme_minimal(base_size = 14) +
  scale_colour_manual(values = region_colours)



```

```{r Tier 3B Reference Comparison}
# Fit GAMs
gam_model <- gam(Value ~ s(Length, by = region) + s(Locname.Fishbase, bs = "re") + s(Sample.Year, k = 10) + region,
                 data = recent_data,
                 method = "REML")

# Null GAM to test for significance of interaction term
gam_null <- gam(Value ~ s(Length) + region,
                data = recent_data,
                method = "REML")

interaction_test <- anova(gam_null, gam_model, test = "Chisq")
interaction_p <- interaction_test$`Pr(>Chi)`[2]
interaction_p_val <- signif(interaction_p, 3)

gam.check(gam_model)

par(mar=c(3,3,3,0),mfrow=c(1,2))
acf(resid(gam_model), lag.max = 36, main = "ACF")
pacf(resid(gam_model), lag.max = 36, main = "pACF")

plot(gam_model)

summary(gam_model)
```

<caption>(#tab:tier-3b-summary-table) Predicted `r contaminant_label` concentrations (µg/g) in `r params$species`, estimated using generalized additive models (GAMs) across the observed range of fish lengths. Values represent model-predicted means with 95% confidence intervals. Asterisks (\*) indicate size classes where the AOC differs significantly (p \< 0.05) from reference.</caption>

```{r tier-3b-summary-table, echo=FALSE}

generate_gam_prediction_table <- function(gam_model, recent_data, params) {
  library(dplyr)
  library(glue)
  library(tidyr)
  library(emmeans)

  # 1. Round size range to nearest 5 cm (outward)
  len_min <- floor(min(recent_data$Length, na.rm = TRUE) / 5) * 5
  len_max <- ceiling(max(recent_data$Length, na.rm = TRUE) / 5) * 5
  length_seq <- seq(len_min, len_max, by = 5)

  # 2. Generate prediction grid
  newdata_grid <- expand.grid(Length = length_seq, region = region_levels)


  # 3. Predict with SE from GAM
  pred_out <- predict(gam_model, newdata = newdata_grid, se.fit = TRUE)
  pred_df <- newdata_grid %>%
    mutate(
      fit = pred_out$fit,
      se = pred_out$se.fit,
      lower = fit - 1.96 * se,
      upper = fit + 1.96 * se
    )

  # 4. Significance test using emmeans at each length
  em <- emmeans(gam_model, ~ region | Length, at = list(Length = length_seq), type = "response")
  em_contrasts <- contrast(em, method = "pairwise", adjust = "none") %>% as.data.frame()

  # 5. Extract lengths where AOC is significantly different
  sig_lengths <- em_contrasts %>%
    filter(grepl("AOC", contrast) & p.value < 0.05) %>%
    pull(Length) %>%
    unique()

  # 6. Add prediction label + asterisk where significant
  pred_df <- pred_df %>%
    mutate(
      sig_star = ifelse(region == "AOC" & Length %in% sig_lengths, "*", ""),
      length_label = paste0(Length, " cm"),
      pred_str = glue("{round(fit, 2)} ({round(lower, 2)}–{round(upper, 2)}){sig_star}")
    )

  # 7. Wide format output: Region × Length (fixed sizes)
  output_table <- pred_df %>%
    select(region, length_label, pred_str) %>%
    pivot_wider(
      names_from = length_label,
      values_from = pred_str
    ) %>%
    rename(Region = region)

  return(output_table)
}



if (interaction_p < 0.05) {
  gam_table <- generate_gam_prediction_table(gam_model, recent_data, params)

  reactable(
    gam_table,
    bordered = TRUE,
    highlight = TRUE,
    striped = TRUE,
    defaultColDef = colDef(align = "center"),
    columns = list(
      Region = colDef(name = "Region", sticky = "left", style = list(fontWeight = "bold"))
    )
  )
}



```

```{r tier-3b-summary-text}

if (interaction_p < 0.05) {
  # Table was generated, so use the same lengths
  len_min <- floor(min(recent_data$Length, na.rm = TRUE) / 5) * 5
  len_max <- ceiling(max(recent_data$Length, na.rm = TRUE) / 5) * 5
  length_seq <- seq(len_min, len_max, by = 5)

  em <- emmeans(gam_model, ~ region | Length, at = list(Length = length_seq), type = "response")
  em_contrasts <- contrast(em, method = "pairwise", adjust = "none") %>% as.data.frame()

  sig_lengths <- em_contrasts %>%
    filter(grepl("AOC", contrast) & p.value < 0.05) %>%
    pull(Length) %>%
    unique()

  sig_count <- length(sig_lengths)
  total_lengths <- length(length_seq)

  t3b_summary_text <- glue::glue(
    "Generalized Additive Models (GAMs) were used to model the relationship between fish length and {contaminant_label} concentrations for {params$species} in the AOC and reference site(s) ({paste(region_levels[-1], collapse = ', ')}). ",
    "The length-contaminant relationship differed significantly between sites (interaction test p = {interaction_p_val}). ",
    "Model-predicted concentrations were calculated across the observed size range (every 5 cm from {min(length_seq)} to {max(length_seq)} cm) and are presented in Table \\@ref(tab:tier-3b-summary-table). ",
    "Asterisks (*) denote sizes where the AOC differs significantly (p < 0.05) from one or more reference site(s); this occurred at {sig_count} of {total_lengths} lengths evaluated."
  )
  
  decision_summary_3b <- em_contrasts %>%
    filter(grepl("AOC", contrast)) %>%
    mutate(higher_aoc = estimate > 0 & p.value < 0.05)

  total_bins <- n_distinct(decision_summary_3b$Length)
  n_fail_b <- sum(decision_summary_3b$higher_aoc, na.rm = TRUE)
  n_pass_b <- total_bins - n_fail_b

  tier3b_decision_text <- glue::glue(
    "Of {total_bins} fish lengths evaluated, {n_fail_b} had significantly higher predicted {contaminant_label} concentrations in the AOC compared to reference (p < 0.05). "
    
  )
 tier3b_overall =  if (n_fail_b > total_bins/2) "Fail" else "Pass"

} else {
  # Use marginal means from the reduced model
  em_summary <- summary(emmeans(gam_null, ~ region))
  em_aoc <- em_summary %>% filter(region == "AOC")
  em_refs <- em_summary %>% filter(region != "AOC")

  ref_texts <- em_refs %>%
    mutate(
      out = glue("{region}: {round(emmean, 2)} {unit_label} (95% CI: {round(lower.CL, 2)}–{round(upper.CL, 2)})")
    ) %>%
    pull(out) %>%
    paste(collapse = "; ")

  t3b_summary_text <- glue::glue(
    "Generalized Additive Models (GAMs) were used to model the relationship between fish length and {contaminant_label} concentrations for {params$species} in the AOC and reference region(s) ({paste(region_levels[-1], collapse = ', ')}). ",
    "The length-contaminant relationship did not differ significantly between sites (interaction test p = {interaction_p_val}). ",
    "Marginal mean concentrations were estimated across the full size distribution using a reduced model with shared smooths. ",
    "The estimated marginal mean {contaminant_label} concentration in the AOC was {round(em_aoc$emmean, 2)} {unit_label} (95% CI: {round(em_aoc$lower.CL, 2)}–{round(em_aoc$upper.CL, 2)}). ",
    "Estimated values for reference site(s): {ref_texts}."
  )
  
  tier3b_decision_text <- if (em_aoc$emmean > max(em_refs$emmean)) {
    glue("The AOC has a higher marginal mean concentration than reference, suggesting elevated contamination in this species compared to reference.")
  } else {
    glue("The AOC does not exceed reference in marginal mean concentration, suggesting comparable contamination levels in this species.")
  }
  
  tier3b_overall <- if (em_aoc$emmean > max(em_refs$emmean)) "Fail" else "Pass"
}



# Construct the summary metric string
if (interaction_p < 0.05) {
  tier3b_metric_text <- glue(
    "Significantly different smooth terms (p = {signif(interaction_p)}); ",
    "{n_fail_b}/{total_bins} lengths with higher AOC concentration"
  )
} else {
  tier3b_metric_text <- glue(
    "No significant difference in smooth terms (p = {round(interaction_p, 3)}); ",
    "AOC marginal mean {if (tier3b_overall == 'Fail') 'higher' else 'not higher'} than reference"
  )
}



tier3b_summary_tibble <- tibble(
  Tier = "Tier 3B",
  Description = "Contaminant Concentration vs Reference",
  Metric = tier3b_metric_text,
  Outcome = tier3b_overall
)


```

<br />

```{r tier3b-summary, results='asis'}
cat(t3b_summary_text)

```

```{r tier-3b-decision-table, results='asis'}
cat(tier3b_decision_text)

```

<br />

**Tier 3B Overall Decision:** `r tier3b_overall`

<br />

## Tier 3C: Temporal Trends

Temporal trends in `r params$species` for `r contaminant_label` were assessed using linear models across all years of data. An initial model tested for a significant interaction between fish length and sampling year. If an interaction was detected, trends were evaluated separately for the three most frequently sampled length bins. Otherwise, annual mean contaminant concentrations were used. Trends help support Tier 3C by assessing if contaminant levels have declined, remained stable, or increased over time.

```{r temp-trends-setup, message = FALSE}
trend_results <- plot_temporal_trends(include_reference = FALSE, include_aoc = TRUE)


temp_trends_caption <- {
  # Interaction phrase
  interaction_phrase <- if (trend_results$interaction_sig) {
    bin_str <- glue::glue_collapse(trend_results$top_bins, sep = ", ", last = " and ")
    glue::glue(
      "Temporal {contaminant_label} concentration ({unit_label}) trends for {params$species}, ",
      "grouped by region and top 3 sampled size bins. ",
      "The most frequently sampled size bins were {bin_str} cm. ",
      "Lines represent linear model fits."
    )
  } else {
    glue::glue(
      "Temporal {contaminant_label} concentration ({unit_label}) trends for {params$species}, ",
      "grouped by region. Lines represent linear model fits to annual mean concentrations."
    )
  }

  # Reference to which regions are included
  included_regions <- c()
  if ("AOC" %in% trend_results$model_stats$region) included_regions <- c(included_regions, "AOC")
  if (ref_1 %in% trend_results$model_stats$region) included_regions <- c(included_regions, ref_1)
  if (!params$combine_ref && ref_2 != "none" && ref_2 %in% trend_results$model_stats$region) {
    included_regions <- c(included_regions, ref_2)
  }

  region_phrase <- glue::glue("Results shown for: {glue::glue_collapse(included_regions, sep = ', ', last = ' and ')}.")

  glue::glue("(ref:temp-trends-fig) {interaction_phrase} {region_phrase}")
}
```

`r temp_trends_caption`

```{r temp-trends-figure, fig.cap='(ref:temp-trends-fig)', fig.width=6,fig.height=4, message=FALSE}
print(trend_results$plot)

interaction_fmt <- ifelse(is.na(trend_results$interaction_p), "NA",
                          ifelse(trend_results$interaction_p < 0.001, "< 0.001", signif(trend_results$interaction_p, 3)))

temporal_text <- {
  # Interaction explanation
  interaction_fmt <- ifelse(
    is.na(trend_results$interaction_p),
    "NA",
    ifelse(trend_results$interaction_p < 0.001, "< 0.001", signif(trend_results$interaction_p, 3))
  )

  interaction_line <- if (trend_results$interaction_sig) {
    top_bins_str <- str_sort(trend_results$top_bins, numeric = TRUE) %>%
  glue::glue_collapse(sep = ", ", last = " and ")
    glue::glue(
      "A linear model detected a significant interaction between year and fish length (p = {interaction_fmt}). ",
      "Trends were therefore assessed separately for the top 3 most sampled length bins: {bin_str} cm."
    )
  } else {
    glue::glue(
      "No significant interaction was detected between year and fish length (p = {interaction_fmt}). ",
      "Trends were assessed using mean contaminant concentrations per year."
    )
  }

  # R² summary
  r2_line <- if ("AOC" %in% trend_results$model_stats$region &&
                 nrow(trend_results$model_stats) == 1) {
    aoc_r2 <- trend_results$model_stats %>% filter(region == "AOC") %>% pull(r2)
    glue::glue("Overall model R² = {round(aoc_r2, 2)}.")
  } else {
    r2_str <- trend_results$model_stats %>%
      mutate(str = glue::glue("**{region}** = {round(r2, 2)}")) %>%
      pull(str) %>%
      glue::glue_collapse(sep = ", ")
    glue::glue("Overall model R² by region: {r2_str}.")
  }

  # Combine both parts
  glue::glue("{interaction_line} {r2_line}")
}



```

```{r tier3c-result-text}
if (trend_results$interaction_sig) {
  # ======= Interaction significant — summarize by bin =======
  tier3c_summary <- trend_results$trend_bins %>%
    filter(region == "AOC") %>%
    count(trend, name = "n") %>%
    mutate(trend = factor(trend, levels = c("Declining", "Stable", "Increasing"))) %>%
    arrange(trend)

  format_trend_phrase <- function(n, trend) {
    verb <- ifelse(n == 1, "was", "were")
    bin_word <- ifelse(n == 1, "bin", "bins")
    glue::glue("{n} {bin_word} {verb} {tolower(trend)}")
  }

  trend_sentences <- pmap_chr(
  list(n = tier3c_summary$n, trend = tier3c_summary$trend),
  format_trend_phrase
) %>%
  glue::glue_collapse(sep = ", ", last = ", and ")


  tier3c_result_text <- glue::glue(
    "In the AOC, {trend_sentences}. These trends were assessed using annual linear models by length bin."
  )

  tier3c_overall <- if (any(tier3c_summary$trend == "Increasing")) "Fail" else "Pass"

} else {
  # ======= No interaction — summarize overall slope =======
  aoc_slope <- trend_results$model_stats %>%
    filter(region == "AOC") %>%
    pull(slope)

  direction <- case_when(
    aoc_slope > 0 ~ "increasing",
    aoc_slope < 0 ~ "decreasing",
    TRUE ~ "stable"
  )

  tier3c_result_text <- glue::glue(
    "In the AOC, {contaminant_label} concentrations were {direction} over time. ",
    "This trend was assessed using an annual linear model of mean concentration."
  )

  tier3c_overall <- if (aoc_slope > 0) "Fail" else "Pass"
}


tier3c_metric_text <- if (trend_results$interaction_sig) {
  # Interaction model — summarize trend direction per bin
  trend_counts <- trend_results$trend_bins %>%
    filter(region == "AOC") %>%
    count(trend) %>%
    tidyr::complete(trend = c("Declining", "Stable", "Increasing"), fill = list(n = 0))

  glue::glue(
    "Significant length × year interaction (p = {interaction_fmt}); ",
    "{trend_counts$n[trend_counts$trend == 'Declining']} declining, ",
    "{trend_counts$n[trend_counts$trend == 'Stable']} stable, ",
    "{trend_counts$n[trend_counts$trend == 'Increasing']} increasing bins"
  )
} else {
  # No interaction — summarize slope direction
  aoc_slope <- trend_results$model_stats %>%
    filter(region == "AOC") %>%
    pull(slope)

  direction <- case_when(
    aoc_slope > 0 ~ "increasing",
    aoc_slope < 0 ~ "decreasing",
    TRUE ~ "stable"
  )

  glue::glue(
    "No significant interaction (p = {interaction_fmt}); ",
    "overall AOC trend: {direction}"
  )
}


tier3c_summary_tibble <- tibble(
  Tier = "Tier 3C",
  Description = "Temporal Trends",
  Metric = tier3c_metric_text,
  Outcome = tier3c_overall
)
```

```{r temporal-text-printout, results='asis'}

cat(temporal_text)
cat(tier3c_result_text)

```

**Tier 3C Overall Decision:** `r tier3c_overall`

# Final summary

```{r}
tier3_summary_table <- bind_rows(
  tier3a_summary_tibble,
  tier3b_summary_tibble,
  tier3c_summary_tibble
)

knitr::kable(tier3_summary_table, caption = "Summary of Tier 3 Evidence Lines and Outcomes")


```
