# Results

## Species of Interest and Restrictive Consumption Thresholds {#results-species-of-interest}

The following species were identified as species of interest, with the respective restrictive consumption thresholds (see Section \@ref(methods-survey)):
 
(ref:threshold-table) Combined fish consumption survey results for the Niagara River AOC. Summary information from NPCA, MCFN, MNO, and SN were compiled to determine species of interest and consumption rates. Species are listed with the percent of survey respondents that eat them, the average consumption rate weighted by average portion size (in meals/month, rounded up to nearest advisory category), and the average consumption rate weighted by average portion size (in meals/month, rounded up to nearest advisory category). Bolded and asterisked species are consumed by at least 5% of respondents, and are therefore considered species of interest.

```{r threshold-table-cap, echo=FALSE, results='asis'}
is_html <- knitr::is_html_output()
  if (is_html){
tbl_caption("threshold-table")
  }
```

```{r threshold-table, tab.cap='(ref:threshold-table)', echo = FALSE}
survey_df = read.csv(here("Data", "survey_results_nr.csv"))


survey_df <- survey_df %>%
  mutate(
    pcteat = as.numeric(pcteat),
    Species = ifelse(pcteat >= 5, paste0(Species, "*"), Species)
  )

# Build flextable
survey_ft <- flextable(survey_df) %>%
  set_header_labels(
    Species = "Species",
    pcteat  = "Percent Eaten by Respondents",
    average = "Weighted Average Consumption Rate (meals/month)",
    top5 = "Weighted Consumption Threshold (meals/month), Top 5% of Respondents"
  ) %>%
  bold(i = ~ pcteat >= 5, j = "Species", bold = TRUE) %>%  # bold species meeting threshold
  colformat_num(
    j = "pcteat",
    suffix = "%"
  ) %>%
  autofit()

survey_ft
```

"Other" species included lake sturgeon, lake whitefish, and other unidentified fishes. Lake whitefish will be considered as a species of interest for this assessment. Lake sturgeon does not have official consumption advice, and will therefore not be included in this assessment.