# Results: Measure 1

## Tier 1 Assessment for the `r AOC_full` AOC {#results-tier1}

```{r t1-setup, echo = F}
# Document setup
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, messages = FALSE)
```

Tier 1 assessment compares the official MECP consumption advisory information for each fish of interest to the desired consumption rate (in meals per month) as indicated by community survey. Table \@ref(tab:tier-1-table) summarizes the MECP advisory values for fish of interest in the `r AOC_full` AOC for both General and Sensitive, compared to each species' community desired consumption rate. Measure 1 assessment is limited to the Lake Ontario (Eastern Basin) advisory zone, which includes the Bay of Quinte.


(ref:tier-1-table-lo) Results of Tier 1 assessment for Lake Ontario (Eastern Basin). The consumption advisory for each species is compared to its desired consumption rate, as determined by community survey. Numbers in cells indicate advisory levels for a species, size class, and population type (General/Sensitive) in the AOC. Red cells indicate an advisory below the "Restrictive" meals/month threshold (indicated by community survey), and green cells indicate advisory levels equal to or above this threshold. The contaminants driving the advisories ("adv cause") are displayed below each cell.

```{r tier-1-table-lo, tab.cap='(ref:tier-1-table-lo)'}
# Load contaminant mappings
contaminant_shapes <- mappings$shapes
contaminant_colours <- mappings$colours

# Render the t1 table using the pared-down RDS table provided from analysis script
t1_table_lo = render_t1_table_any(
  df = t1_df_lo,
  contaminant_shapes = contaminant_shapes,
  contaminant_colours = contaminant_colours,
  generate_shape_fn = generate_shape,
  shape_size = 8,
  force = "flextable",
  table_height = "1000px",
  show_legend = TRUE
)


t1_table_lo
```

(ref:tier-1-pass-fail-lo) Results by species of Tier 1 assessment for Lake Ontario (Eastern Basin). The consumption advisory for each species is compared to its desired consumption rate, as determined by community survey. Species for which the consumption advisory is equal to or above the desired consumption rate pass Tier 1 and are excluded from further analysis; otherwise the species fails and moves forward to Tier 2.

```{r tier-1-pass-fail-lo, results = 'asis', tab.cap='(ref:tier-1-pass-fail-lo)'}
  
t1_pass = report_pass_fail_species(flags_lo, tier = "t1", output = "table")


```




