---
title: "NR T2 Reference Sites"
always_allow_html: true
output:
  html_document:
    theme: yeti
    toc: false
  pdf_document:
    toc: false
params:
  AOC: '42577857'
  AOC_shorthand: "NR-UR"
  reference_sites:
  - '44007727'
  - '43557712'
  - '43497905'
  - '43117922'
  - '43397900'
  - '43517847'
  - '43527851'
  - '44397535'
  - '43557717'
  - '43527856'
  restrict_threshold: 8
  thr_path: "Data/consumption_threshold_nr.csv"
  interest_species:
  - Black Crappie
  - Channel Catfish
  - Freshwater Drum
  - Largemouth Bass
  - Northern Pike
  - Rock Bass
  - Smallmouth Bass
  - Walleye
  - White Perch
  - Yellow Perch
  - Chinook Salmon
  - Lake Whitefish
  - Rainbow Smelt
  - Rainbow Trout
  - Lake Trout
  - Brown Trout
  - Coho Salmon
  - Silver Bass
---

<style type="text/css">
.main-container {
  max-width: 1800px;
  margin-left: auto;
  margin-right: auto;
}
</style>

# Upper Niagara River
### Reference Sites - Open Water Erie
```{r, echo = F, include = F}
knitr::opts_chunk$set(echo = F)
options(width = 1800)
source("Scripts/setup.R")
```


```{r}
heatmap_df = cons_data %>%
  filter(specname %in% params$interest_species) 
erie_sites = c("Lake Erie", "Wheatley Harbour (Lake Erie)", "Rondeau Bay (Lake Erie)")
ontario_open = c("Lake Ontario (Eastern Basin)", "Lake Ontario (Western Basin)")
open_sites = c("Lake Erie", ontario_open)

ontario_sites = c("Lake Ontario (Western Basin)", "Lake Ontario (Eastern Basin)", "Jordan Harbour (Lake Ontario)", "Frenchman Bay (Lake Ontario)", "Whitby Harbour (Lake Ontario)", "McLaughlin Bay (Lake Ontario)", "Cataraqui Bay (Lake Ontario)", "East Lake", "West Lake")
all_sites = c(erie_sites, ontario_sites)

get_waterbody_ids(cons_data,all_sites)



```


```{r}
library(mapview)

library(sf)

map_df = heatmap_df %>%
  filter(guide_locname_eng %in% all_sites)

map_df <- map_df %>%
  mutate(
    lat_dd = latitude / 10000,
    lon_dd = longitude / 10000
  ) %>%
  mutate(lon_dd = ifelse(lon_dd > 0, -lon_dd, lon_dd)) %>%
  distinct(guide_locname_eng, lat_dd, lon_dd)

sites_sf <- map_df %>%
  st_as_sf(coords = c("lon_dd", "lat_dd"), crs = 4326)

sites_map = st_transform(sites_sf, 3857)
xy = st_coordinates(sites_map)

label_df = sites_map %>%
  st_drop_geometry() %>%
  bind_cols(as.data.frame(xy))

bb = st_bbox(sites_map)

pad = 120000

xlim <- c(bb["xmin"] - pad, bb["xmax"] + pad)
ylim <- c(bb["ymin"] - pad, bb["ymax"] + pad)

library(ggspatial)
library(ggrepel)

ggplot() +
  annotation_map_tile(type = "osm", zoomin = 0) +
  geom_sf(data = sites_map, size = 3, color = "red") +
  geom_text_repel(
    data = label_df,
    aes(x = X, y = Y, label = guide_locname_eng),
    size = 3,
    min.segment.length = 0,
    max.overlaps = Inf
  ) +
  annotation_scale(location = "bl", width_hint = 0.3) +
  annotation_north_arrow(
    location = "br",
    which_north = "true",
    style = north_arrow_fancy_orienteering
  ) +
  coord_sf(crs = 3857,
           xlim = xlim,
           ylim = ylim,expand = FALSE) +
  theme_minimal() +
  labs(x = NULL, y = NULL) +
  theme(axis.text = element_blank())
```

#### Reference sites used:

- Lake Erie

```{r, out.width="100%"}

# Heatmaps
make_heatmap_table(df = heatmap_df, length_levels = length_levels, aoc_site_name = "Niagara River (Upper reach)", sites = "Lake Erie")

```



### Reference Sites - All Erie

#### Reference sites used:
`r knitr::asis_output(paste0("- ", erie_sites, collapse = "\n"))`

```{r}

# Heatmaps
make_heatmap_table(df = heatmap_df, length_levels = length_levels, aoc_site_name = "Niagara River (Upper reach)", sites = erie_sites)

```



### Reference Sites - Open Water Erie + Ontario

#### Reference sites used:
`r knitr::asis_output(paste0("- ", open_sites, collapse = "\n"))`

```{r}
# Heatmaps
make_heatmap_table(df = heatmap_df, length_levels = length_levels, aoc_site_name = "Niagara River (Upper reach)", sites = open_sites)

```



### Reference Sites - All Erie + Ontario


#### Reference sites used:

`r knitr::asis_output(paste0("- ", all_sites, collapse = "\n"))`


```{r}

 
# Heatmaps
make_heatmap_table(df = heatmap_df, length_levels = length_levels, aoc_site_name = "Niagara River (Upper reach)", sites = all_sites)


```



# Lower Niagara River



### Reference Sites - Open Water Ontario

#### Reference sites used:

`r knitr::asis_output(paste0("- ", ontario_open, collapse = "\n"))`


```{r}
# Heatmaps
make_heatmap_table(df = heatmap_df, length_levels = length_levels, aoc_site_name = "Niagara River (Lower reach)", sites = ontario_open)

```



### Reference Sites - All Ontario


#### Reference sites used:

`r knitr::asis_output(paste0("- ", ontario_sites, collapse = "\n"))`


```{r}
# Heatmaps
make_heatmap_table(df = heatmap_df, length_levels = length_levels, aoc_site_name = "Niagara River (Lower reach)", sites = ontario_sites)

```


### Reference Sites - Open Water Ontario + Erie


#### Reference sites used:

`r knitr::asis_output(paste0("- ", open_sites, collapse = "\n"))`


```{r}
# Heatmaps
make_heatmap_table(df = heatmap_df, length_levels = length_levels, aoc_site_name = "Niagara River (Lower reach)", sites = open_sites)

```


### Reference Sites - All Ontario + Erie

#### Reference sites used:

`r knitr::asis_output(paste0("- ", all_sites, collapse = "\n"))`

```{r}
# Heatmaps
make_heatmap_table(df = heatmap_df, length_levels = length_levels, aoc_site_name = "Niagara River (Lower reach)", sites = all_sites)

```

