---
title: "NR T2 Reference Sites"
always_allow_html: true
output:
  html_document:
    theme: yeti
    toc: false
  pdf_document:
    toc: false
params:
  AOC: '42577857'
  AOC_shorthand: "NR-UR"
  reference_sites:
  - '44007727'
  - '43557712'
  - '43497905'
  - '43117922'
  - '43397900'
  - '43517847'
  - '43527851'
  - '44397535'
  - '43557717'
  - '43527856'
  restrict_threshold: 8
  thr_path: "Data/consumption_threshold_nr.csv"
  interest_species:
  - Black Crappie
  - Channel Catfish
  - Freshwater Drum
  - Largemouth Bass
  - Northern Pike
  - Rock Bass
  - Smallmouth Bass
  - Walleye
  - White Perch
  - Yellow Perch
  - Chinook Salmon
  - Lake Whitefish
  - Rainbow Smelt
  - Rainbow Trout
  - Lake Trout
  - Brown Trout
  - Coho Salmon
  - Silver Bass
---

<style type="text/css">
.main-container {
  max-width: 1800px;
  margin-left: auto;
  margin-right: auto;
}
</style>

# Upper Niagara River
### Reference Sites - Open Water Erie
```{r, echo = F, include = F}
knitr::opts_chunk$set(echo = F)
options(width = 1800)
source("Scripts/setup.R")
```

```{r, out.width="100%"}
heatmap_df = cons_data %>%
  filter(specname %in% params$interest_species) 
# Heatmaps
make_heatmap_table(df = heatmap_df, length_levels = length_levels, aoc_site_name = "Niagara River (Upper reach)", sites = "Lake Erie")

```
Lake Erie


### Reference Sites - All Erie
```{r}
erie_sites = c("Lake Erie", "Wheatley Harbour (Lake Erie)", "Rondeau Bay (Lake Erie)")

# Heatmaps
make_heatmap_table(df = heatmap_df, length_levels = length_levels, aoc_site_name = "Niagara River (Upper reach)", sites = erie_sites)

```

`r knitr::asis_output(paste0("- ", erie_sites, collapse = "\n"))`

### Reference Sites - Open Water Erie + Ontario
```{r}
ontario_open = c("Lake Ontario (Eastern Basin)", "Lake Ontario (Western Basin)")
open_sites = c("Lake Erie", ontario_open)



# Heatmaps
make_heatmap_table(df = heatmap_df, length_levels = length_levels, aoc_site_name = "Niagara River (Upper reach)", sites = open_sites)

```

`r knitr::asis_output(paste0("- ", open_sites, collapse = "\n"))`

### Reference Sites - All Erie + Ontario
```{r}
ontario_sites = c("Lake Ontario (Western Basin)", "Lake Ontario (Eastern Basin)", "Jordan Harbour (Lake Ontario)", "Port Dalhousie (Lake Ontario)", "Frenchman Bay (Lake Ontario)", "Whitby Harbour (Lake Ontario)", "McLaughlin Bay (Lake Ontario)", "Cataraqui Bay (Lake Ontario)")
all_sites = c(erie_sites, ontario_sites)
 
# Heatmaps
make_heatmap_table(df = heatmap_df, length_levels = length_levels, aoc_site_name = "Niagara River (Upper reach)", sites = all_sites)


```

`r knitr::asis_output(paste0("- ", all_sites, collapse = "\n"))`

# Lower Niagara River

### Reference Sites - Open Water Ontario
```{r}
# Heatmaps
make_heatmap_table(df = heatmap_df, length_levels = length_levels, aoc_site_name = "Niagara River (Lower reach)", sites = ontario_open)

```

`r knitr::asis_output(paste0("- ", ontario_open, collapse = "\n"))`

### Reference Sites - All Ontario
```{r}
# Heatmaps
make_heatmap_table(df = heatmap_df, length_levels = length_levels, aoc_site_name = "Niagara River (Lower reach)", sites = ontario_sites)

```

`r knitr::asis_output(paste0("- ", ontario_sites, collapse = "\n"))`

### Reference Sites - Open Water Ontario + Erie
```{r}
# Heatmaps
make_heatmap_table(df = heatmap_df, length_levels = length_levels, aoc_site_name = "Niagara River (Lower reach)", sites = open_sites)

```

`r knitr::asis_output(paste0("- ", open_sites, collapse = "\n"))`

### Reference Sites - All Ontario + Erie
```{r}
# Heatmaps
make_heatmap_table(df = heatmap_df, length_levels = length_levels, aoc_site_name = "Niagara River (Lower reach)", sites = all_sites)

```

`r knitr::asis_output(paste0("- ", all_sites, collapse = "\n"))`