# Results

## Species of Interest and Desired Consumption Rate {#results-species-of-interest}

The following species were identified as species of interest, with the respective desired consumption rates (see Section \@ref(methods-survey)):
 
(ref:threshold-table) Community survey results in the St. Lawrence River (Cornwall/Akwesasne) AOC. Species are listed with the percent of survey respondents that eat them, the current consumption rate weighted by average portion size (in meals/month, rounded up to nearest advisory category), and the desired consumption rate weighted by average portion size (in meals/month, rounded up to nearest advisory category). Bolded and asterisked species are consumed by at least 5% of respondents, and are therefore considered species of interest.

```{r threshold-table-cap, echo=FALSE, results='asis'}
is_html <- knitr::is_html_output()
  if (is_html){
tbl_caption("threshold-table")
  }
```

```{r threshold-table, tab.cap='(ref:threshold-table)', echo = FALSE}
survey_df = read.csv(here("Data", "survey_results_slr.csv"))


survey_df <- survey_df %>%
  mutate(
    pcteat = as.numeric(pcteat),
    Species = ifelse(pcteat >= 5, paste0(Species, "*"), Species)
  )

# Build flextable
survey_ft <- flextable(survey_df) %>%
  set_header_labels(
    Species = "Species",
    pcteat  = "Percent Eaten by Respondents",
    current = "Current Weighted Consumption Rate (meals/month)",
    desired = "Desired Weighted Consumption Rate (meals/month)"
  ) %>%
  bold(i = ~ pcteat >= 5, j = "Species", bold = TRUE) %>%  # bold species meeting threshold
  colformat_num(
    j = "pcteat",
    suffix = "%"
  ) %>%
  autofit()

survey_ft
```

Of the identified species of interest, neither black crappie or chinook salmon have official consumption advice in the AOC, and will therefore be excluded from the tiered assessment.