---
title: "NR BUI 1 Tier 3 Analysis"
author: "Aaron Dain"
always_allow_html: true
output:
  bookdown::html_document2:
    theme: yeti
    css: fullwidth.css
    toc: true
    toc_float: true
  html_document:
    toc: true
    df_print: paged
  pdf_document:
    toc: true
    latex_engine: wkhtmltopdf
  html_notebook:
    self_contained: true
  bookdown::pdf_document2:
    toc: true
params:
  AOC_name: Lake Ontario 1a - Upper Niagara River
  species: Rainbow Trout
  contaminant: PCBs
  aoc_shapefile: null
  raw_csv: "Data/Great Lakes Data to Ken 2024-12 PCB-Hg(Data).csv"
  ref_1: Lake Ontario
  ref_2: Lake Erie
  combine_ref: TRUE
  add_AOC: 
  - "Lake Ontario 1a"
  exclude_site:
  - Hamilton Harbour
  - Trent River
  - Niagara River
  - Toronto Waterfront
  - Detroit River
  - Bay of Quinte
  target_sr: 4326
  restrict_threshold: 8
  exclude_missing: FALSE
  save_artifacts: TRUE
  out_dir_base: "Output/Full-NPCA-Report/Derived/NR/Tier3"
---

```{r self-ref, eval = F}
# Self reference of species of interest for boq that failed t2

# Brown Bullhead	❌
# Chinook Salmon	❌ ⚠️
# Lake Whitefish	— ⚠️
# Northern Pike	❌
# Rainbow Smelt	❌ ⚠️
# Smallmouth Bass	❌
# Walleye	❌


```


```{r Setup, include = FALSE}

out_dir = here::here(params$out_dir_base)


# Loading setup script
library(here)
source(here::here("Scripts", "setup.R"))

raw_data = read.csv(params$raw_csv)

# -------- filenames & folders ----------
# where the book will read from (relative to book root)
out_dir_base   <- params$out_dir_base                       # "Output/Full-SLR-Report/Derived/SLR/Tier3"
run_id         <- paste0(params$species, "__", params$contaminant)
out_dir_rel    <- file.path(out_dir_base, run_id)           # e.g., "Output/.../Walleye__MERCURY"
out_dir_abs    <- here::here(out_dir_rel)                   # absolute for writing files

fs::dir_create(file.path(out_dir_abs, "figs"))
fs::dir_create(file.path(out_dir_abs, "tables"))
fs::dir_create(file.path(out_dir_abs, "text"))
fs::dir_create(file.path(out_dir_abs, "models"))
fs::dir_create(file.path(out_dir_abs, "html"))




# toggle
SAVE <- isTRUE(params$save_artifacts)

# central registry for this knit
ARTIFACTS <- list(
  params = params,
  figs   = list(),
  tables = list(),
  html   = list(),
  text   = list(),
  data   = list(),
  models = list(),
  decisions = list(),
  summary = NULL
)


# Restrict threshold override
thr_path <- "Data/consumption_threshold_nr.csv"
thr_df <- if (file.exists(thr_path)) {readr::read_csv(thr_path, show_col_types = FALSE)} else {tibble::tibble()}

new_thr = thr_df %>%
  filter(Species == params$species) %>%
  pull(as.integer(threshold))

# Save new threshold to artifacts
ARTIFACTS$params$restrict_threshold = as.integer(new_thr)

# -------- small savers -----------
# returns "figs/<name>.png"
save_gg <- function(plot, name, w=7, h=5, dpi=300) {
  if (!SAVE || is.null(plot)) return(NULL)
  rel <- file.path("figs", glue("{name}.png"))
  abs <- file.path(out_dir_abs, rel)
  ggplot2::ggsave(abs, plot = plot, width = w, height = h, dpi = dpi, bg = "white")
  rel
}

# returns "html/<name>.html"
save_widget <- function(widget, name) {
  if (!SAVE || is.null(widget)) return(NULL)
  rel <- file.path("html", glue("{name}.html"))
  abs <- file.path(out_dir_abs, rel)
  htmlwidgets::saveWidget(widget, file = abs, selfcontained = TRUE)
  rel
}

# returns list(csv = "tables/<name>.csv", rds = "tables/<name>.rds")
save_df <- function(df, name) {
  if (!SAVE || is.null(df)) return(NULL)
  rel_csv <- file.path("tables", glue("{name}.csv"))
  rel_rds <- file.path("tables", glue("{name}.rds"))
  readr::write_csv(df, file.path(out_dir_abs, rel_csv))
  saveRDS(df, file.path(out_dir_abs, rel_rds))
  list(csv = rel_csv, rds = rel_rds)
}

save_kable_html <- function(kbl, name) {
  if (!SAVE || is.null(kbl)) return(NULL)
  rel <- file.path("html", glue("{name}.html"))
  kableExtra::save_kable(kbl, file.path(out_dir_abs, rel))
  rel
}

save_text <- function(txt, name) {
  if (!SAVE || is.null(txt)) return(NULL)
  rel <- file.path("text", glue("{name}.txt"))
  if (length(txt) > 1) txt <- paste(txt, collapse = "\n")
  writeLines(txt, file.path(out_dir_abs, rel), useBytes = TRUE)
  rel
}

save_model <- function(obj, name) {
  if (!SAVE || is.null(obj)) return(NULL)
  rel <- file.path("models", glue("{name}.rds"))
  saveRDS(obj, file.path(out_dir_abs, rel))
  rel
}

```



```{r Options, include = FALSE}
# Setup Webshot 2 to browswer to allow tables to be rendered in PDF. 
# IMPORTANT: Please set this up correctly - the path to your browser application.
Sys.setenv(CHROMOTE_CHROME = "C:/Program Files/BraveSoftware/Brave-Browser/Application/brave.exe")


# Makes contaminant label title-capitalized
contaminant_label <- if (params$contaminant == "MERCURY") {
str_to_title(params$contaminant)
} else {params$contaminant}

# Get unit label
get_unit_label <- function(contaminant) {
  contaminant_upper <- toupper(contaminant)
  if (contaminant_upper == "MERCURY") {
    return("µg/g")
  } else if (contaminant_upper == "PCBs") {
    return("ng/g")
  } else if (contaminant_upper == "TEQWH4") {
    return("pg/g") 
  } else {
    return("units")
  }
}

# Creates label for units dynamically usign get_unit_label function
unit_label <- get_unit_label(params$contaminant)



```

```{r Dataset-Filtering, include = FALSE}


# Filter raw data by species and contaminant
filtered_data <- raw_data %>%
  filter(Specname == params$species,
         Contaminant == params$contaminant) %>%
  mutate(region = NA_character_)

has_coords <- !is.na(filtered_data$Longitude.Decimal) & !is.na(filtered_data$Latitude.Decimal)


# Dynamically assign ref_1 and ref_2 based on whether combining reference regions
if (params$combine_ref) {
  ref_1 <- "Reference"
  ref_2 <- "none"
  region_levels <- c("AOC", ref_1)
  ref_label <- "Reference"
  ref_regions_to_match <- c(params$ref_1, params$ref_2)
} else {
  ref_1 <- params$ref_1
  ref_2 <- params$ref_2
  region_levels <- c("AOC", ref_1, if (ref_2 != "none") ref_2 else NULL)
  ref_label <- if (ref_2 != "none") paste(ref_1, "and", ref_2) else ref_1
  ref_regions_to_match <- c(ref_1, if (ref_2 != "none") ref_2 else NULL)
}

# Load shapefile for AOC region and reproject

# Check if AOC shapefile exists
use_aoc_shapefile <- !is.null(params$aoc_shapefile) && file.exists(params$aoc_shapefile) 
# If there is shapefile for AOC. Is ignored if not
if (use_aoc_shapefile) {
  aoc_shp <- st_read(params$aoc_shapefile) %>% st_transform(crs = params$target_sr)

  filtered_sf <- filtered_data[has_coords, ] %>%
    st_as_sf(coords = c("Longitude.Decimal", "Latitude.Decimal"), crs = params$target_sr) %>%
    st_transform(crs = st_crs(aoc_shp)) %>%
    st_make_valid()

  aoc_match <- st_intersects(filtered_sf, st_make_valid(aoc_shp), sparse = FALSE)[, 1]
  filtered_data$region[has_coords][aoc_match] <- "AOC"
}


# Assigns region, "AOC" and reference names (option to combine ref_1 and ref_2 as "Reference")
filtered_data <- filtered_data %>%
  mutate(
    region = case_when(
      region == "AOC" ~ "AOC",
      is.na(region) & str_detect(Locname.Fishbase, regex(str_c(params$add_AOC, collapse = "|"), ignore_case = TRUE)) ~ "AOC",

      params$combine_ref & is.na(region) & (
        str_detect(Locname.Fishbase, fixed(params$ref_1, ignore_case = TRUE)) |
        str_detect(Locname.Fishbase, fixed(params$ref_2, ignore_case = TRUE))
      ) ~ "Reference",

      !params$combine_ref & is.na(region) & str_detect(Locname.Fishbase, fixed(params$ref_1, ignore_case = TRUE)) ~ params$ref_1,

      !params$combine_ref & is.na(region) & 
        !is.null(params$ref_2) & 
        params$ref_2 != "none" &
        str_detect(Locname.Fishbase, fixed(params$ref_2, ignore_case = TRUE)) ~ params$ref_2,

      TRUE ~ region
    )
  ) %>%
  filter(
    !is.na(region),
    region == "AOC" | !str_detect(Locname.Fishbase, str_c(params$exclude_site, collapse = "|"))
  ) %>%
  mutate(region = factor(region))

# Full analysis data - for temporal trends
full_data <- if (params$exclude_missing) {
  filtered_data %>% filter(!is.na(Longitude.Decimal), !is.na(Latitude.Decimal))
} else {
  filtered_data
}

# Mapping data (recent + coordinates only)
map_data <- filtered_data %>%
  filter(Sample.Year >= 2014,
         !is.na(Longitude.Decimal),
         !is.na(Latitude.Decimal))

missing_coords <- filtered_data %>%
  filter(Sample.Year >= 2014,
         is.na(Longitude.Decimal) | is.na(Latitude.Decimal))
missing_sites <- unique(missing_coords$Locname.Fishbase) %>% sort()

recent_data <- filtered_data %>%
  filter(Sample.Year >= 2014) %>%
  { if (params$exclude_missing) filter(., !is.na(Longitude.Decimal) & !is.na(Latitude.Decimal)) else . }


# Assigning colours to regions
region_colours <- setNames(
  c("red", "cyan", "magenta")[seq_along(region_levels)],
  region_levels
)



```


```{r data-sufficiency}
# --- Tier 3 pre-check: sufficient AOC data for this species? ------------------


# (1) Filter to this species (if a Species/species column exists) and AOC region


fd_aoc <- filtered_data %>%
  filter(region == "AOC")

# (2) Count unique years and total observations in AOC
n_years <- fd_aoc %>% distinct(Sample.Year) %>% nrow()
n_obs   <- nrow(fd_aoc)

# (3) Define sufficiency rule
sufficient <- (n_years >= 5) && (n_obs >= 30)

if (!sufficient) {
  # Build an "Insufficient Data" summary for Tiers 3A–3C
  insuf_metric <- paste0(
    "Available years: ", n_years,
    "; observations: ", n_obs,
    " (require ≥5 years and ≥30 observations in the AOC)"
  )

  insuf_summary <- tibble(
    Species     = params$species,
    Tier        = c("Tier 3A", "Tier 3B", "Tier 3C"),
    Description = c(
      "Virtual Advisory",
      "Contaminant Concentration vs Reference",
      "Temporal Trends"
    ),
    Metric  = insuf_metric,
    Outcome = "Insufficient Data"
  )

  # Ensure ARTIFACTS list & species slot exist, then write summary
  if (!exists("ARTIFACTS", inherits = TRUE) || is.null(ARTIFACTS)) ARTIFACTS <- list()

  ARTIFACTS$tables$summary <- save_df(insuf_summary, "tier3_summary_table")

  ARTIFACTS$summary <- insuf_summary

  # Write a compact index.rds pointing to everything you saved
  idx_file <- file.path(out_dir, "index.rds")
  # … after populating ARTIFACTS …
  saveRDS(ARTIFACTS, file.path(out_dir_abs, "index.rds"))
  writeLines(c(
    glue("species: {params$species}"),
    glue("contaminant: {params$contaminant}"),
    glue("output_dir_rel: {out_dir_rel}"),       # useful for debugging
    glue("index_rds_rel: index.rds")
  ), file.path(out_dir_abs, "INDEX.txt"))

  message(
    sprintf(
      "⚠️ Insufficient data for %s in %s: %d years, %d observations. ",
      params$species, "AOC", n_years, n_obs
    ),
    "Wrote Insufficient Data summary for Tiers 3A–3C and stopping Tier 3."
  )

  # Stop further analysis/knitting for this run
  knitr::knit_exit()
  stop()
}

# If sufficient, continue with Tier 3...


```

# Tier 3 Analysis for `r contaminant_label` in `r params$species` for the `r params$AOC_name` AOC

## Reference Sites

<!-- (ref:map) Map showing AOC and reference site locations. -->

<!-- ```{r map-setup, eval = F} -->
<!-- # Function for consistent aspect ratio for map -->
<!-- adjust_bbox_to_aspect <- function(bbox, aspect_ratio = 10/6, crs = 3857) { -->
<!--   # Validate input -->
<!--   if (anyNA(bbox)) { -->
<!--     warning("adjust_bbox_to_aspect: bbox contains NA values.") -->
<!--     return(NULL) -->
<!--   } -->

<!--   # Extract numeric coordinates -->
<!--   xmin <- bbox["xmin"] -->
<!--   xmax <- bbox["xmax"] -->
<!--   ymin <- bbox["ymin"] -->
<!--   ymax <- bbox["ymax"] -->

<!--   width <- xmax - xmin -->
<!--   height <- ymax - ymin -->

<!--   if (width <= 0 || height <= 0) { -->
<!--     warning("adjust_bbox_to_aspect: invalid bbox dimensions.") -->
<!--     return(NULL) -->
<!--   } -->

<!--   current_ratio <- width / height -->
<!--   center_x <- (xmin + xmax) / 2 -->
<!--   center_y <- (ymin + ymax) / 2 -->

<!--   if (current_ratio > aspect_ratio) { -->
<!--     # Too wide → expand height -->
<!--     new_height <- width / aspect_ratio -->
<!--     new_ymin <- center_y - new_height / 2 -->
<!--     new_ymax <- center_y + new_height / 2 -->
<!--     new_bbox <- c(xmin = xmin, ymin = new_ymin, xmax = xmax, ymax = new_ymax) -->
<!--   } else { -->
<!--     # Too tall → expand width -->
<!--     new_width <- height * aspect_ratio -->
<!--     new_xmin <- center_x - new_width / 2 -->
<!--     new_xmax <- center_x + new_width / 2 -->
<!--     new_bbox <- c(xmin = new_xmin, ymin = ymin, xmax = new_xmax, ymax = ymax) -->
<!--   } -->

<!--   # Ensure new_bbox has correct names and is a named numeric vector -->
<!--   new_bbox_named <- setNames(new_bbox, c("xmin", "ymin", "xmax", "ymax")) -->

<!--   # Now construct bbox object -->
<!--   st_bbox(new_bbox_named, crs = st_crs(crs)) -->
<!-- } -->



<!-- ``` -->

<!-- ```{r map-creation, eval=F, message=FALSE, include=FALSE} -->

<!-- library(maptiles) -->
<!-- library(ggspatial) -->

<!-- # Convert to sf object and project -->
<!-- sites_sf <- st_as_sf(map_data, coords = c("Longitude.Decimal", "Latitude.Decimal"), crs = params$target_sr) -->
<!-- sites_proj <- st_transform(sites_sf, 3857)  # Web Mercator -->

<!-- # Main map bounding box -->
<!-- bbox_buffered <- st_bbox(sites_proj) -->
<!-- bbox_adjusted <- adjust_bbox_to_aspect(bbox_buffered, aspect_ratio = 10/6, crs = 3857) -->
<!-- bbox_sfc <- st_as_sfc(bbox_adjusted + c(-40000, -40000, 40000, 40000)) -->

<!-- # Get basemap -->
<!-- basemap <- get_tiles(bbox_sfc, provider = "Esri.WorldShadedRelief", crop = TRUE, zoom = 10) -->
<!-- bbox_vec <- st_bbox(basemap) -->

<!-- # -------- AOC inset logic (only if shapefile is being used) -------- -->
<!-- aoc_sites <- sites_proj[sites_proj$region == "AOC", ] -->
<!-- has_inset <- nrow(aoc_sites) > 0 -->

<!-- if (isTRUE(has_inset)) { -->
<!--   aoc_bbox <- st_bbox(aoc_sites) + c(-10000, -10000, 10000, 10000) -->
<!--   aoc_bbox_sfc <- st_as_sfc(aoc_bbox) -->
<!--   aoc_basemap <- get_tiles(aoc_bbox_sfc, provider = "Esri.WorldShadedRelief", crop = TRUE, zoom = 12) -->

<!--   aoc_inset_sites <- sites_proj[st_within(sites_proj, aoc_bbox_sfc, sparse = FALSE), ] -->
<!-- } -->


<!-- ``` -->

<!-- ```{r map, fig.cap='(ref:map)', message = FALSE,eval=FALSE} -->
<!-- aspect_ratio = 6/10 -->

<!-- map_plot <- NULL -->

<!-- # Only draw if we have coords -->
<!-- if (nrow(map_data) > 0) { -->
<!--   sites_sf  <- sf::st_as_sf(map_data, coords = c("Longitude.Decimal","Latitude.Decimal"), crs = params$target_sr) -->
<!--   sites_proj <- sf::st_transform(sites_sf, 3857) -->

<!--   bb <- safe_bbox(sites_proj) -->
<!--   if (!is.null(bb)) { -->
<!--     bbox_adjusted <- adjust_bbox_to_aspect(bb, aspect_ratio = 10/6, crs = 3857) -->
<!--     bbox_sfc <- st_as_sfc(bbox_adjusted + c(-40000, -40000, 40000, 40000)) -->
<!--     basemap <- try(maptiles::get_tiles(bbox_sfc, provider = "Esri.WorldShadedRelief", crop = TRUE, zoom = 10), silent = TRUE) -->

<!--     aoc_sites <- sites_proj[sites_proj$region == "AOC", ] -->
<!--     has_inset <- nrow(aoc_sites) > 0 -->

<!--     main_map <- ggplot() + -->
<!--       layer_spatial(basemap) + -->
<!--       geom_sf(data = sites_proj, aes(color = region), size = 3.5, shape = 21, fill = "white", stroke = 1) + -->
<!--       coord_sf(xlim = c(st_bbox(basemap)["xmin"], st_bbox(basemap)["xmax"]), -->
<!--                ylim = c(st_bbox(basemap)["ymin"], st_bbox(basemap)["ymax"]), -->
<!--                expand = FALSE, clip = "on", datum = NA) + -->
<!--       ggspatial::annotation_scale(location = "bl", width_hint = 0.2) + -->
<!--       labs(color = "Region") + -->
<!--       theme_void(base_size = 12) + -->
<!--       theme( -->
<!--         legend.justification = c(1,0), -->
<!--         legend.position = c(.98,.02), -->
<!--         legend.box.background = element_rect(fill = "#EEFFEF", color = "black"), -->
<!--         legend.box.margin = unit(c(2,2,2,2),  "pt"), -->
<!--         plot.margin = unit(c(0,0,0,0), "pt") -->
<!--       ) + scale_color_manual(values = region_colours) -->

<!--     if (has_inset) { -->
<!--       aoc_bbox <- st_bbox(aoc_sites) + c(-10000, -10000, 10000, 10000) -->
<!--       aoc_bbox_sfc <- st_as_sfc(aoc_bbox) -->
<!--       aoc_basemap <- try(maptiles::get_tiles(aoc_bbox_sfc, provider = "Esri.WorldShadedRelief", crop = TRUE, zoom = 12), silent = TRUE) -->
<!--       aoc_inset_sites <- sites_proj[st_within(sites_proj, aoc_bbox_sfc, sparse = FALSE), ] -->

<!--       inset_map <- ggplot() + -->
<!--         layer_spatial(aoc_basemap) + -->
<!--         geom_sf(data = aoc_inset_sites, aes(color = region), size = 2.5, shape = 21, fill = "white", stroke = 0.8) + -->
<!--         ggspatial::annotation_scale(location = "bl", width_hint = 0.3) + -->
<!--         coord_sf(xlim = c(aoc_bbox["xmin"], aoc_bbox["xmax"]), -->
<!--                  ylim = c(aoc_bbox["ymin"], aoc_bbox["ymax"]), expand = FALSE) + -->
<!--         scale_colour_manual(values = region_colours) + -->
<!--         theme_void() + -->
<!--         theme(aspect.ratio = NULL, panel.border = element_rect(color = "black", fill = NA), legend.position = "none") -->

<!--       main_map <- main_map +  geom_sf(data = aoc_bbox_sfc, fill = NA, color = "red", linewidth = 0.5) -->
<!--       map_plot <- cowplot::ggdraw(main_map) + -->
<!--         cowplot::draw_plot(inset_map, x = -0.04, y = 0.6, width = 0.4, height = 0.35) + -->
<!--         theme(plot.margin = unit(c(0,0,0,0), "pt")) -->
<!--     } else { -->
<!--       map_plot <- main_map -->
<!--     } -->
<!--   } -->
<!-- } -->

<!-- if (!is.null(map_plot)) { -->
<!--   print(map_plot) -->
<!--   ARTIFACTS$figs$map <- save_gg(map_plot, "map") -->
<!-- } else { -->
<!--   message("Map not created (no coordinates).") -->
<!-- } -->
<!-- ``` -->


```{r Display-Reference-Sites, include = FALSE}
# Step 1: Ensure distinct region-site combinations
ref_site_listing <- recent_data %>%

  distinct(region, site = Locname.Fishbase) %>%
  arrange(region, site)

# Step 2: Collapse to one entry per region
ref_sites_md <- ref_site_listing %>%
  group_by(region) %>%
  reframe(
    markdown = paste0("#### ", unique(region), " {-}", "\n", paste0("- ", site, collapse = "\n")),
    .groups = "drop"
  ) %>%
  distinct(markdown) %>%  # <-- Remove duplicates if grouped improperly upstream
  pull(markdown) %>%
  paste(collapse = "\n\n")


ARTIFACTS$text$reference_sites <- save_text(ref_sites_md, "reference_sites")

```

<br />

The following reference sites were used in the Tier 3 analysis of `r contaminant_label` in `r params$species` for the `r params$AOC_names` AOC:

<br />

`r ref_sites_md`

<br />

```{r report-missing-sites-message, results = "asis"}

# Print out missing location sites

if (length(missing_sites) > 0) {
  if (params$exclude_missing) {
    cat(glue::glue("### Note\n\n {length(missing_sites)} sites with missing coordinates were excluded **from the analysis**.\n\n"))
  } else {
    cat(glue::glue("### Note\n\nRetained {length(missing_sites)} sites with missing coordinates **for analysis**, but excluded them from maps.\n\n"))
  }
  
  cat("The following sites were affected:\n\n")
  cat(paste0("- ", missing_sites, collapse = "\n"))
}

# Create a single string and save
missing_msg <- if (length(missing_sites) > 0) {
  if (params$exclude_missing) {
    paste0("Note: ", length(missing_sites), " sites with missing coordinates were excluded from the analysis.\n\n",
           "Affected sites:\n- ", paste(missing_sites, collapse = "\n- "))
  } else {
    paste0("Note: Retained ", length(missing_sites), " sites with missing coordinates for analysis (excluded from maps).\n\n",
           "Affected sites:\n- ", paste(missing_sites, collapse = "\n- "))
  }
} else ""
ARTIFACTS$text$missing_sites <- save_text(missing_msg, "missing_sites")

```

## Tier 3A: Virtual Advisory for `r contaminant_label` in `r params$species`

Virtual advisory values were generated using by modelling the relationship between fish length and contaminant concentration for `r params$species` and `r contaminant_label`. For each region, the function fits a log-log linear model (lm(log(Value) \~ log(Length)) to evaluate whether a significant relationship exists (p ≤ 0.05). If so, predicted concentrations are generated across a continuous length range, and then averaged within fixed 5 cm bins. Advisory levels are assigned using MECP-defined contaminant- and population-specific (General/Sensitive) thresholds to the mean predicted concentrations. If no significant relationship is detected, advisory levels are based on the mean concentration of all observations within each bin.

```{r Generate virtual advisories, include = FALSE}
virtual_advisory = generate_virtual_advisories(
  df = recent_data,
  group_var = "region",
  contaminant = params$contaminant
)

```

(ref:Virtual-advisory-visualization) Virtual `r contaminant_label` advisories in `r params$species`. Predicted concentrations (curve) and observed values (points) by region using power regression model, compared to MECP advisory limits for `r params$contaminant` (dashed lines).

```{r Virtual-advisory-visualization, fig.cap='(ref:Virtual-advisory-visualization)'}
# Extract predictions and original observations
pred_df <- virtual_advisory$predictions
obs_df <- recent_data  
contaminant <- params$contaminant

y_label <- if (contaminant == "MERCURY") {
  "Mercury Concentration (µg/g)"
} else if (contaminant == "PCBs") {
  "PCBs Concentration (ng/g)"
} else if (contaminant == "TEQWH4"){
  "Dioxin/Furans TEQ (TEQWH, pg/g)"
} else {
  paste(contaminant_label, "(unknown units)")
}

get_thresholds <- function(contaminant = "MERCURY", population = "Sensitive") {
  if (contaminant == "MERCURY") {
    if (population == "Sensitive") {
      return(tibble(
        meals = c(16, 12, 8, 4, 0),
        conc = c(0.06, 0.12, 0.16, 0.25, 0.50)
      ))
    } else {
      return(tibble(
        meals = c(32, 16, 12, 8, 4, 2, 0),
        conc = c(0.15, 0.15, 0.3, 0.4, 0.6, 1.2, 1.8)
      ))
    }
  } else if (contaminant == "PCBs") {
    return(tibble(
      meals = c(32, 16, 12, 8, 4, 2, 1, 0),
      conc = c(26,  26, 53, 70, 105, 211, 422, 844)
    ))
  
} else if (contaminant == "TEQWH4") {
    return(tibble(
    meals = c(32, 16, 12, 8, 4, 2, 1, 0),
    conc = c(0.7, 0.7, 1.3, 1.8, 2.7, 5.4, 10.8, 21.6)
      ))
  } else {
    
  return(NULL)
}
}

thresholds_df <- get_thresholds(contaminant, population = "Sensitive")


# Plot with overlaid observations and advisory thresholds
va_plot =  ggplot(pred_df, aes(x = Length, y = predicted, color = region)) +
  geom_point(data = obs_df, aes(x = Length, y = Value, color = region), size = 1.5, alpha = 0.7) +
  geom_line(linewidth = 1.2) +
  geom_hline(data = thresholds_df, aes(yintercept = conc), linetype = "dashed", color = "black") +
  geom_text(data = thresholds_df, aes(x = Inf, y = conc, label = paste0(meals, " meals/mo")),
            hjust = 1.1, vjust = -0.2, color = "black", size = 3, inherit.aes = FALSE) +
  scale_colour_manual(values = region_colours) +
  labs(
    x = "Length (cm)",
    y = y_label,
    color = "Region"
  ) +
  theme_classic(base_size = 13)
va_plot
ARTIFACTS$figs$t3a_virtual <- save_gg(va_plot, "t3a_virtual_advisory")

```

<br />

```{r Model-summary-creation, include=FALSE}
model_summary <- virtual_advisory$model_info %>%
  mutate(
    summary_text = pmap_chr(list(lm, correlated, pval, r2), function(mod, sig, pval, r2) {
      coef_vals <- coef(mod)
      a <- signif(exp(coef_vals[1]), 3)
      b <- round(coef_vals[2], 3)
      pval_fmt <- ifelse(pval < 0.001, "< 0.001", formatC(pval, format = "e", digits = 2))
      r2_fmt <- round(r2, 3)

      sig_text <- if (sig) {
        "There is a statistically significant relationship between fish length and contaminant concentration."
      } else {
        "There is no significant relationship between fish length and contaminant concentration. Bin-level means were used instead."
      }

      model_text <- glue("The power model used was: concentration = {a} × Length<sup>{b}</sup> (R² = {r2_fmt}, p = {pval_fmt}).")

      paste0(sig_text, " ", model_text)
    })
  ) %>%
  select(region, summary_text)

model_summary_txt <- paste0(
  "Model Summary by Region\n\n",
  paste0("**", model_summary$region, "**: ", model_summary$summary_text, collapse = "\n\n")
)
ARTIFACTS$text$t3a_model_summary <- save_text(model_summary_txt, "t3a_model_summary")

```

<br />

```{r model-summary-printout, results='asis'}
# display model summaries
cat(paste0("**Model Summary by Region:** {-} \n\n ",
           paste0("**", model_summary$region, "**: ", model_summary$summary_text, collapse = "\n")))

```

<caption>(#tab:virtual-advisory-table) Virtual advisory levels for `r contaminant_label` in `r params$species` across regions. Values represent meals/month recommendations. Red cells indicate restrictive virtual advisory falling below the desired consumption rate of `r new_thr` meals/month; green cells indicate unrestrictive virtual advisory equal to or above the desired consumption rate.</caption>

```{r virtual-advisory-table, results='asis', warning = TRUE}

predictions_binned <- virtual_advisory$predictions %>%
  mutate(
    length_bin = cut(
      Length,
      breaks = c(seq(15, 75, by = 5), Inf),
      labels = c(paste(seq(15, 70, by = 5), seq(20, 75, by = 5), sep = "-"), "75+"),
      right = FALSE
    )
  ) %>%
  filter(!is.na(length_bin)) %>%
  group_by(region, length_bin) %>%
  summarise(
    mean_conc = mean(predicted, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  mutate(
  meals_general   = assign_advisory(mean_conc, "General", params$contaminant),
  meals_sensitive = assign_advisory(mean_conc, "Sensitive", params$contaminant)
)





color_advisory_cell <- function(value, threshold = new_thr) {
  if (is.na(value)) return("")
  if (value < threshold) {
    return(list(background = adv_palette$fail, color = "white", fontWeight = "bold"))
  } else {
    return(list(background = adv_palette$pass, color = "black", fontWeight = "bold"))
  }
}



virtual_long <- predictions_binned %>%
  pivot_longer(cols = c(meals_general, meals_sensitive),
               names_to = "Population",
               names_prefix = "meals_",
               values_to = "Advisory") %>%
  mutate(
    Population = case_when(
      Population == "general" ~ "General",
      Population == "sensitive" ~ "Sensitive"
    )
  )%>%
  filter(
    toupper(params$contaminant) == "MERCURY" | Population == "General"
  )


virtual_wide <- virtual_long %>%
  select(Region = region, Population, length_bin, Advisory) %>%
  pivot_wider(names_from = length_bin, values_from = Advisory)


virtual_adv_table <- reactable(
  virtual_wide,
  bordered = TRUE,
  highlight = TRUE,
  striped = TRUE,
  defaultColDef = colDef(
    style = function(value) {
      if (is.numeric(value) && !is.na(value)) {
        color_advisory_cell(value)
      } else {
        NULL
      }
    },
    align = "center"
  ),
  columns = list(
    Region = colDef(name = "Region", sticky = "left", style = list(fontWeight = "bold")),
    Population = colDef(name = "Population", sticky = "left", style = list(fontWeight = "bold"))
  )
)


virtual_adv_table


ARTIFACTS$tables$t3a_virtual_advisory <- save_df(virtual_wide, "t3a_virtual_advisory_table")
ARTIFACTS$html$t3a_virtual_advisory  <- save_widget(virtual_adv_table, "t3a_virtual_advisory_table")

```

```{r tier-3a-decision}
# Decision making


# Step 3: Is AOC more restrictive than reference? Split AOC and reference tables
aoc_advisories <- virtual_long %>%
  filter(region == "AOC") %>%
  select(length_bin, Population, Advisory) %>%
  rename(AOC_Advisory = Advisory)

reference_advisories <- virtual_long %>%
  filter(region != "AOC") %>%
  group_by(length_bin, Population) %>%
  summarise(Ref_Advisory = median(Advisory, na.rm = TRUE), .groups = "drop")  # Median across references - important if >1 reference group


# Join and evaluate
tier3a_joined <- aoc_advisories %>%
  left_join(reference_advisories, by = c("length_bin", "Population")) %>%
  mutate(
    restrict_pass = AOC_Advisory >= new_thr,
    restrict_fail = !restrict_pass,
    more_restrictive_than_ref = AOC_Advisory < Ref_Advisory,
    final_decision = case_when(
      restrict_pass ~ "Pass",
      restrict_fail & !more_restrictive_than_ref ~ "Pass",
      restrict_fail & more_restrictive_than_ref ~ "Fail",
      TRUE ~ "Unknown"
    )
  )


tier3a_summary <- tier3a_joined %>%
  group_by(Population) %>%
  summarise(
    n_bins = n(),
    n_below_thresh = sum(restrict_fail),
    n_below_ref = sum(restrict_fail & more_restrictive_than_ref),
    n_pass = n_bins - n_below_ref,
    pct_pass = round(100 * n_pass / n_bins, 1),
    .groups = "drop"
  ) %>%
  mutate(
    decision = if_else(n_below_ref == 0, "Pass", "Fail")
  )



tier3a_text <- tier3a_summary %>%
  mutate(
    summary = glue::glue(
      "**{Population}** population: Of {n_bins} size bins, {n_below_thresh} were below the desired consumption rate of {params$restrict_threshold} meals/month. ",
      "Of these, {n_below_ref} were more restrictive than reference."
    )
  ) %>%
  pull(summary) %>%
  glue::glue_collapse(sep = " ")

tier3a_overall <- if (any(tier3a_summary$decision == "Fail")) "Unsupportive" else "Supportive"


tier3a_summary_tibble <- tibble(
  Tier = "Tier 3A",
  Description = "Virtual Advisory",
  Metric = glue_collapse(glue("{tier3a_summary$Population}: {tier3a_summary$n_below_ref}/{tier3a_summary$n_bins} size bins restrictive or more restrictive than reference"), sep = "; "),
  Outcome = tier3a_overall
)
```

```{r tier-3a-decision-table, results='asis'}
cat(tier3a_text)

ARTIFACTS$text$t3a_decision_text <- save_text(tier3a_text, "t3a_decision_text")
ARTIFACTS$decisions$t3a_overall   <- tier3a_overall


```

<br />

**Tier 3A Overall Decision:** `r tier3a_overall`

<br />

## Tier 3B: Comparison of contaminant concentrations in `r params$species` in AOC to reference

To assess differences in fish contamination between the AOC and reference sites, contaminant–length relationships were modeled using Generalized Additive Models (GAMs), allowing for non-linear associations. For each priority contaminant and indicator species, we first tested whether the shape of the concentration–length relationship differed significantly between AOC and reference fish. Where models supported statistically similar trends (non-significant interaction), we compared size-adjusted contaminant levels using estimated marginal means. Where shapes differed significantly, contaminant concentration values were compared by GAM-predicted values at standardized lengths (e.g., 45cm) commonly consumed for each fish species. A threshold of α = 0.05 was used to evaluate statistical significance throughout.

(ref:Tier-3B-Figure) `r contaminant_label` concentration in `r params$species` by length in the AOC compared to reference sites. Points are observed contaminant values, curves represent GAM model fit (with random effects dropped for simplicity, fitted across the 1st to 99th percentile of data to ensure reasonable support), ribbons represent 95% confidence interval.



```{r Tier 3B-Reference-Comparison, include = FALSE}
# Model data, factorizing inputs
gam_df = recent_data %>%
  mutate(Waterbody.Code = factor(Waterbody.Code), Sample.Year = factor(Sample.Year))


# Fit gam robustly given possible single-level factors
library(mgcv)
library(dplyr)
library(glue)

fit_gam <- function(gam_df, k_len = 10, method = "REML", verbose = TRUE) {
  # Ensure factors & drop unused
  gam_df <- gam_df %>%
    mutate(
      region = factor(region),
      Waterbody.Code = factor(Waterbody.Code)
    ) %>%
    droplevels() %>%
    mutate(
      Value = if_else(Value <= 0, 0.001, Value) # GAM breaks on values <= 0, this ensures it works when below detection limits
    )

  n_region <- nlevels(gam_df$region)
  n_site   <- nlevels(gam_df$Waterbody.Code)

  # Base term always present
  base <- glue("log(Value) ~ s(Length, k = {k_len})")

  # Optional terms based on cardinality
  has_region <- n_region >= 2
  has_siteRE <- n_site   >= 2

  # FULL model: s(Length) + fs(Length, region) [if >=2] + s(Waterbody.Code, re) [if >=2]
  full_terms <- base
  if (has_region) full_terms <- glue("log(Value) ~ s(Length, region, bs = 'fs', k = {k_len})")
  if (has_siteRE) full_terms <- paste0(full_terms, " + s(Waterbody.Code, bs = 're')")
  f_full <- as.formula(full_terms)

  # NULL model: s(Length) + region main effect [if >=2] + s(Waterbody.Code, re) [if >=2]
  null_terms <- base
  if (has_region) null_terms <- paste0(null_terms, " + region")
  if (has_siteRE) null_terms <- paste0(null_terms, " + s(Waterbody.Code, bs = 're')")
  f_null <- as.formula(null_terms)

  if (verbose) {
    message("Region levels: ", n_region, " | Site levels: ", n_site)
    message("NULL: ", deparse(f_null))
    message("FULL: ", deparse(f_full))
  }

  # Fit models
  gam_null <- gam(f_null, data = gam_df, method = method)
  gam_full <- gam(f_full, data = gam_df, method = method)

  # If fs term couldn't be included (has_region == FALSE), no interaction test is meaningful
  if (!has_region) {
    if (verbose) message("Skipping interaction test: only one region level present.")
    return(list(
      gam_null = gam_null,
      gam_full = gam_full,
      interaction_p = NA_real_,
      interaction_msg = "Not tested (single region level)."
    ))
  }

  # If formulas end up identical (e.g., both lack fs term), skip test
  if (identical(f_null, f_full)) {
    if (verbose) message("Skipping interaction test: NULL and FULL are identical.")
    return(list(
      gam_null = gam_null,
      gam_full = gam_full,
      interaction_p = pval,
      interaction_msg = "Not tested (null and full identical)."
    ))
  }

  # Run LR test on smooth structure
  atab <- anova(gam_null, gam_full, test = "Chisq")
  # try to pull the right cell; fallbacks cover weird table layouts
  pcol <- suppressWarnings(as.numeric(atab$`Pr(>Chi)`))
  pval <- if (length(pcol) >= 2) pcol[2] else if (length(pcol) >= 1) pcol[length(pcol)] else NA_real_
  if (!is.finite(pval)) pval <- NA_real_

  list(
    gam_null = gam_null,
    gam_full = gam_full,
    interaction_p = pval,
    interaction_msg = "Tested (NULL: s(Length)+region+RE; FULL: + fs(Length,region))."
  )
}


# Fit GAMs
res <- fit_gam(gam_df, k_len = 5, method = "REML", verbose = TRUE)

gam_model <- res$gam_full
gam_null <- res$gam_null
interaction_p <- res$interaction_p
interaction_p_val <- if (!is.na(interaction_p)) signif(interaction_p, 3) else NA_character_ 

ARTIFACTS$models$gam_full <- save_model(gam_model, "t3b_gam_full")
ARTIFACTS$models$gam_null <- save_model(gam_null, "t3b_gam_null")

```

```{r gam-validation, eval = FALSE}

# Run this chunk to evaluate GAMs

gam.check(gam_model)

par(mar=c(3,3,3,0),mfrow=c(1,2))
acf(resid(gam_model), lag.max = 36, main = "ACF")
pacf(resid(gam_model), lag.max = 36, main = "pACF")

plot(gam_model)

summary(gam_model)


gam_valid = validate_gam(gam_model, data = gam_df)




knitr::kable(
  gam_valid %>%
    transmute(
      family, link, n, r2 = round(r2, 3),
      dev_expl = round(dev_expl, 3),
      acf_flag, pacf_flag, lb_flag,
      concurvity_max = round(concurvity_max, 3), concurvity_flag
    ),
  caption = "GAM validation summary"
)

# Expand the per-term tables when needed (e.g., in collapsible sections)
smooth_terms   <- gam_valid$smooth_table[[1]]
param_terms    <- gam_valid$param_table[[1]]
k_diagnostics  <- gam_valid$k_table[[1]]
lb_diagnostics <- gam_valid$lb_table[[1]]

# Example: flag text you can drop into the report
issue_msgs <- c(
  if (gam_valid$acf_flag)  "Residual ACF shows significant autocorrelation." else NULL,
  if (gam_valid$lb_flag)   "Ljung–Box test indicates autocorrelation at some lags." else NULL,
  if (any(k_diagnostics$k_flag, na.rm = TRUE)) "Some smooths suggest k may be too low (k-index < 1)." else NULL,
  if (gam_valid$concurvity_flag) "High concurvity detected (> 0.90). Consider simplifying terms." else NULL
)
if (length(issue_msgs) == 0) issue_msgs <- "No major validation flags detected."
cat(paste0("**Validation notes:** ", paste(issue_msgs, collapse = " ")))


concurvity(gam_model, full = T)


region_levels <- levels(model.frame(gam_model)$region)
exclude_re <- names(gam_model$smooth)[sapply(gam_model$smooth, function(s) s$bs == "re")]

```

<caption>(#tab:tier-3b-summary-table) Predicted `r contaminant_label` concentrations (µg/g) in `r params$species`, estimated using generalized additive models (GAMs) across the observed range of fish lengths where data is available for both groups. Values represent model-predicted means with 95% confidence intervals. Asterisks (\*) indicate size classes where the AOC differs significantly (p \< 0.05) from reference.</caption>

```{r tier-3b-summary-table, echo=FALSE}

generate_gam_prediction_table <- function(
  gam_model,
  recent_data,
  compare_to = "Reference",
  alpha = 0.05,
  back_transform = TRUE
) {
  library(dplyr); library(glue); library(tidyr)

  mf <- model.frame(gam_model)
  region_levels <- levels(mf$region)

  # ===== Guard 1: need at least 1 region; if only 1, skip sig tests later =====
  has_compare <- (length(region_levels) >= 2) && (compare_to %in% region_levels)
  aoc_name <- if ("AOC" %in% region_levels) "AOC" else setdiff(region_levels, compare_to)[1]

  # ===== 1) Length grid (rounded to 5 cm) =====
  qs <- recent_data %>%
    group_by(region) %>%
    summarise(lo = quantile(Length, .01, na.rm = TRUE),
              hi = quantile(Length, .99, na.rm = TRUE), .groups = "drop")

  lo5 <- ceiling(max(qs$lo, na.rm = TRUE) / 5) * 5
  hi5 <- floor(min(qs$hi, na.rm = TRUE) / 5) * 5
  if (!is.finite(lo5) || !is.finite(hi5) || lo5 > hi5) {
    # Fallback: use overall recent_data range
    lo5 <- ceiling(min(recent_data$Length, na.rm = TRUE) / 5) * 5
    hi5 <- floor(max(recent_data$Length, na.rm = TRUE) / 5) * 5
  }
  length_seq <- seq(lo5, hi5, by = 5)
  if (length(length_seq) == 0L) {
    stop("No valid length sequence could be generated for predictions.")
  }

  # ===== 2) Newdata grid =====
  newdata_grid <- expand.grid(
    Length = length_seq,
    region = region_levels,
    KEEP.OUT.ATTRS = FALSE
  )
  newdata_grid$region <- factor(newdata_grid$region, levels = region_levels)

  if ("Waterbody.Code" %in% names(mf)) {
    if (is.factor(mf$Waterbody.Code)) {
      newdata_grid$Waterbody.Code <- factor(levels(mf$Waterbody.Code)[1],
                                            levels = levels(mf$Waterbody.Code))
    } else newdata_grid$Waterbody.Code <- mf$Waterbody.Code[1]
  }
  if ("Sample.Year" %in% names(mf)) {
    if (is.factor(mf$Sample.Year)) {
      newdata_grid$Sample.Year <- factor(levels(mf$Sample.Year)[1],
                                         levels = levels(mf$Sample.Year))
    } else newdata_grid$Sample.Year <- median(recent_data$Sample.Year, na.rm = TRUE)
  }

  # ===== 3) Predictions excluding RE smooths =====
  re_labels <- vapply(gam_model$smooth,
                      function(s) if (!is.null(s$bs) && s$bs == "re") s$label else NA_character_,
                      character(1))
  re_labels <- stats::na.omit(re_labels)

  Xp   <- predict(gam_model, newdata = newdata_grid, type = "lpmatrix", exclude = re_labels)
  beta <- coef(gam_model)
  V    <- vcov(gam_model, unconditional = FALSE)

  fit   <- as.numeric(Xp %*% beta)
  se    <- sqrt(rowSums((Xp %*% V) * Xp))
  lower <- fit - 1.96 * se
  upper <- fit + 1.96 * se

  pred_df <- cbind(newdata_grid, fit = fit, se = se, lower = lower, upper = upper) |> as.data.frame()

  if (back_transform) {
    pred_df <- pred_df %>%
      mutate(fit_bt = exp(fit), lower_bt = exp(lower), upper_bt = exp(upper))
  } else {
    pred_df <- pred_df %>%
      mutate(fit_bt = fit, lower_bt = lower, upper_bt = upper)
  }

  # ===== 4) AOC vs Reference tests (only if both present) =====
  sig_lengths <- integer(0)
  if (has_compare && !is.na(aoc_name) && (aoc_name %in% region_levels)) {
    row_idx <- function(r, L) which(pred_df$region == r & pred_df$Length == L)[1]
    pvals <- lapply(length_seq, function(L) {
      ia <- row_idx(aoc_name,   L)
      ir <- row_idx(compare_to, L)
      if (is.na(ia) || is.na(ir)) return(data.frame(Length = L, p = NA_real_))
      Xa <- Xp[ia, , drop = FALSE]
      Xr <- Xp[ir, , drop = FALSE]
      d  <- Xa - Xr
      est <- as.numeric(d %*% beta)
      seD <- sqrt(as.numeric(d %*% V %*% t(d)))
      p   <- if (seD > 0) 2 * pnorm(-abs(est / seD)) else NA_real_
      data.frame(Length = L, p = p)
    }) %>% bind_rows()
    sig_lengths <- pvals %>% filter(is.finite(p), p < alpha) %>% pull(Length)
  }

  # ===== 5) Compose display =====
  pred_df <- pred_df %>%
    mutate(
      sig_star = ifelse(region %in% aoc_name & Length %in% sig_lengths, "*", ""),
      length_label = paste0(Length, " cm"),
      pred_str = glue("{round(fit_bt, 2)} ({round(lower_bt, 2)}–{round(upper_bt, 2)}){sig_star}")
    )

  output_table <- pred_df %>%
    select(region, length_label, pred_str) %>%
    pivot_wider(names_from = length_label, values_from = pred_str) %>%
    rename(Region = region)

  attr(output_table, "preds") <- pred_df
  output_table
}


if (isTRUE(interaction_p < 0.05)) {
  gam_table <- generate_gam_prediction_table(
    gam_model, recent_data, compare_to = "Reference", alpha = 0.05, back_transform = TRUE
  )

  gam_reactable =  reactable(
    gam_table,
    bordered = TRUE, highlight = TRUE, striped = TRUE,
    defaultColDef = colDef(align = "center"),
    columns = list(Region = colDef(name = "Region", sticky = "left",
                                   style = list(fontWeight = "bold")))
  )

  length_seq <- sort(unique(attr(gam_table, "preds")$Length))
  
  gam_reactable
}



if (isTRUE(interaction_p < 0.05) && exists("gam_table")) {
  ARTIFACTS$tables$t3b_pred <- save_df(as.data.frame(gam_table), "t3b_pred_table")
  if (exists("gam_reactable")) ARTIFACTS$html$t3b_pred <- save_widget(gam_reactable, "t3b_pred_table")
}

```

```{r Tier-3B-Figure, echo = FALSE, message = FALSE, fig.cap='(ref:Tier-3B-Figure)'}
# --- 1) minimal predictions (dense grid per region; REs dropped) ---
mf <- model.frame(gam_model)

has_region_in_model <- "region" %in% names(mf)
region_levels_model <- if (has_region_in_model) levels(mf$region) else NULL

# fall back to regions in the data if model has no region term
region_levels_data <- sort(unique(as.character(recent_data$region)))
region_levels <- if (!is.null(region_levels_model)) region_levels_model else region_levels_data

# guard: if still empty, bail early
stopifnot(length(region_levels) >= 1)

# length grid
len_seq_raw <- seq(min(recent_data$Length, na.rm = TRUE),
                   max(recent_data$Length, na.rm = TRUE),
                   length.out = 600)

# build newdata grid with a region column no matter what
preds_plot <- tidyr::crossing(
  region = region_levels,
  Length = len_seq_raw
)

# align region type for prediction: factor w/ model levels if model has region term
if (has_region_in_model) {
  preds_plot$region <- factor(preds_plot$region, levels = region_levels_model)
} else {
  # keep as character; predict.gam will ignore unknown cols anyway
  preds_plot$region <- as.character(preds_plot$region)
}

# stubs for RE/other columns if they exist in the model frame
if ("Waterbody.Code" %in% names(mf)) {
  preds_plot$Waterbody.Code <-
    if (is.factor(mf$Waterbody.Code)) factor(levels(mf$Waterbody.Code)[1], levels = levels(mf$Waterbody.Code))
    else mf$Waterbody.Code[1]
}
if ("Sample.Year" %in% names(mf)) {
  preds_plot$Sample.Year <-
    if (is.factor(mf$Sample.Year)) factor(levels(mf$Sample.Year)[1], levels = levels(mf$Sample.Year))
    else stats::median(recent_data$Sample.Year, na.rm = TRUE)
}

# collect labels for random-effect smooths safely
exclude_re <- vapply(gam_model$smooth, function(s) {
  bs <- tryCatch(s$bs, error = function(...) NULL)
  if (!is.null(bs) && identical(bs, "re")) s$label else NA_character_
}, character(1))
exclude_re <- stats::na.omit(exclude_re)

# predictions (model scale -> back-transform)
pr <- predict(gam_model, newdata = preds_plot, type = "link", se.fit = TRUE,
              exclude = exclude_re, unconditional = FALSE)

preds_plot <- preds_plot |>
  dplyr::mutate(
    fit = as.numeric(pr$fit),
    se  = as.numeric(pr$se.fit),
    fit_bt   = exp(fit),
    lower_bt = exp(fit - 1.96*se),
    upper_bt = exp(fit + 1.96*se)
  )

# --- 2) clip to in-range lengths per region (works even if only one region) ---
len_limits <- recent_data |>
  dplyr::group_by(region) |>
  dplyr::summarise(
    Lmin = stats::quantile(Length, 0.01, na.rm = TRUE),
    Lmax = stats::quantile(Length, 0.99, na.rm = TRUE),
    .groups = "drop"
  ) |>
  dplyr::mutate(region = as.character(region))  # ensure same type for join

preds_plot <- preds_plot |>
  dplyr::mutate(region = as.character(region)) |>
  dplyr::inner_join(len_limits, by = "region") |>
  dplyr::filter(Length >= Lmin, Length <= Lmax)

# --- 3) plot (will show one curve if only one region) ---
t3b_plot = ggplot(recent_data, aes(Length, Value, color = region)) +
  geom_point(alpha = 0.6) +
  geom_ribbon(data = preds_plot, aes(ymin = lower_bt, ymax = upper_bt, x = Length, fill = region),
              alpha = 0.15, inherit.aes = FALSE) +
  geom_line(data = preds_plot, aes(y = fit_bt, x = Length, color = region), linewidth = 1.2) +
  labs(
       x = "Length (cm)", y = y_label, color = "Region", fill = "Region") +
  theme_classic(base_size = 14) +
  scale_colour_manual(values = region_colours, guide = "legend") +
  scale_fill_manual(values = region_colours, guide = "legend")

t3b_plot
ARTIFACTS$figs$t3b_gam_plot <- save_gg(t3b_plot, "t3b_gam_plot")

```


```{r tier-3b-summary-text}
lpmatrix_contrasts <- function(gam_model, length_seq, recent_data,
                               aoc_name = "AOC") {
  mf   <- model.frame(gam_model)
  regs <- levels(mf$region)
  refs <- setdiff(regs, aoc_name)
  stopifnot(length(refs) > 0)

  # Prediction grid (region × chosen lengths) + placeholders for RE factors
  newgrid <- expand.grid(region = regs, Length = length_seq, KEEP.OUT.ATTRS = FALSE)
  newgrid$region <- factor(newgrid$region, levels = regs)
  if ("Waterbody.Code" %in% names(mf)) {
    newgrid$Waterbody.Code <- if (is.factor(mf$Waterbody.Code))
      factor(levels(mf$Waterbody.Code)[1], levels = levels(mf$Waterbody.Code))
    else mf$Waterbody.Code[1]
  }
  if ("Sample.Year" %in% names(mf)) {
    newgrid$Sample.Year <- if (is.factor(mf$Sample.Year))
      factor(levels(mf$Sample.Year)[1], levels = levels(mf$Sample.Year))
    else median(recent_data$Sample.Year, na.rm = TRUE)
  }

  # Region-level design (drop all bs='re')
  Xp   <- predict(gam_model, newdata = newgrid, type = "lpmatrix",
                  exclude = exclude_re, unconditional = FALSE)
  beta <- coef(gam_model)
  V    <- vcov(gam_model, unconditional = FALSE)

  idx <- function(r, L) which(newgrid$region == r & newgrid$Length == L)[1]

  # AOC vs each reference at each length (on log scale)
  out <- lapply(length_seq, function(L) {
    ia <- idx(aoc_name, L)
    do.call(rbind, lapply(refs, function(ref) {
      ir <- idx(ref, L)
      d  <- Xp[ia, , drop = FALSE] - Xp[ir, , drop = FALSE]
      est <- as.numeric(d %*% beta)
      se  <- sqrt(as.numeric(d %*% V %*% t(d)))
      data.frame(Length = L, ref = ref, est = est, se = se,
                 z = est / se, p = 2 * pnorm(-abs(est / se)))
    }))
  })
  dplyr::bind_rows(out)
}



if (isTRUE(interaction_p < 0.05)) {
  stopifnot(exists("length_seq"), length(length_seq) > 0)

  contr <- lpmatrix_contrasts(gam_model, length_seq, recent_data, aoc_name = "AOC")

  # Lengths where AOC differs from ANY reference at α=0.05
  sig_lengths <- contr |>
    dplyr::group_by(Length) |>
    dplyr::summarise(any_sig = any(p < 0.05, na.rm = TRUE), .groups = "drop") |>
    dplyr::filter(any_sig) |>
    dplyr::pull(Length)

  sig_count     <- length(sig_lengths)
  total_lengths <- length(length_seq)

  # Count bins where AOC is significantly HIGHER than any reference
  decision_summary_3b <- contr |>
    dplyr::mutate(higher_aoc = est > 0 & p < 0.05)

  total_bins <- dplyr::n_distinct(decision_summary_3b$Length)
  n_fail_b   <- decision_summary_3b |>
                  dplyr::group_by(Length) |>
                  dplyr::summarise(any_higher = any(higher_aoc, na.rm = TRUE), .groups = "drop") |>
                  dplyr::summarise(n = sum(any_higher)) |>
                  dplyr::pull(n)
  n_pass_b <- total_bins - n_fail_b

  # (Optional) ratios on original scale for reporting, if you want them
  contr <- contr |>
    dplyr::mutate(ratio = exp(est),
                  lcl_ratio = exp(est - 1.96 * se),
                  ucl_ratio = exp(est + 1.96 * se))

  # Narrative text (same as your template)
  t3b_summary_text <- glue::glue(
    "Generalized Additive Models (GAMs) were used to model the relationship between fish length and {tolower(contaminant_label)} concentrations for {tolower(params$species)} in the AOC and reference sites. ",
    "The length–contaminant relationship differed significantly between sites (interaction test p = {interaction_p_val}). ",
    "Model-predicted concentrations were calculated across the observed size range for which observations existed for both regions (every 5 cm from {min(length_seq)} to {max(length_seq)} cm).",
  )

  tier3b_decision_text <- glue::glue(
    "Of {total_bins} fish lengths evaluated, {n_fail_b} had significantly higher predicted {tolower(contaminant_label)} concentrations in the AOC compared to reference (p < 0.05). "
  )
  tier3b_overall <- if (n_fail_b > total_bins/2) "Unsupportive" else "Supportive"

} else {

  # Use marginal means from the reduced model
  em_summary <- summary(emmeans(gam_null, ~ region))
  em_aoc <- em_summary %>% filter(region == "AOC")
  em_refs <- em_summary %>% filter(region != "AOC")

  ref_texts <- em_refs %>%
    mutate(
      out = glue("{region}: {round(emmean, 2)} {unit_label} (95% CI: {round(lower.CL, 2)}–{round(upper.CL, 2)})")
    ) %>%
    pull(out) %>%
    paste(collapse = "; ")

  t3b_summary_text <- glue::glue(
    "Generalized Additive Models (GAMs) were used to model the relationship between fish length and {tolower(contaminant_label)} concentrations for {tolower(params$species)} in the AOC and reference regions. ",
    "The length-contaminant relationship did not differ significantly between sites (interaction test p = {interaction_p_val}). ",
    "Marginal mean concentrations were estimated across the full size distribution using a reduced model with shared smooths. ",
    "The estimated marginal mean {contaminant_label} concentration in the AOC was {round(em_aoc$emmean, 2)} {unit_label} (95% CI: {round(em_aoc$lower.CL, 2)}–{round(em_aoc$upper.CL, 2)}). ",
    "Estimated values for reference site(s): {ref_texts}."
  )
  
  tier3b_decision_text <- if (em_aoc$emmean > max(em_refs$emmean)) {
    glue("The AOC has a higher marginal mean concentration than reference, suggesting elevated contamination in this species compared to reference.")
  } else {
    glue("The AOC does not exceed reference in marginal mean concentration, suggesting comparable contamination levels in this species.")
  }
  
  tier3b_overall <- if (em_aoc$emmean > max(em_refs$emmean)) "Unsupportive" else "Supportive"
}




# Construct the summary metric string
if (isTRUE(interaction_p < 0.05)) {
  tier3b_metric_text <- glue(
    "Significantly different smooth terms (p = {signif(interaction_p)}); ",
    "{n_fail_b}/{total_bins} lengths with higher AOC concentration"
  )
} else {
  tier3b_metric_text <- glue(
    "No significant difference in smooth terms (p = {round(interaction_p, 3)}); ",
    "AOC marginal mean {if (tier3b_overall == 'Unsupportive') 'higher' else 'not higher'} than reference"
  )
}



tier3b_summary_tibble <- tibble(
  Tier = "Tier 3B",
  Description = "Contaminant Concentration vs Reference",
  Metric = tier3b_metric_text,
  Outcome = tier3b_overall
)

ARTIFACTS$text$t3b_summary  <- save_text(t3b_summary_text, "t3b_summary_text")
ARTIFACTS$text$t3b_decision <- save_text(tier3b_decision_text, "t3b_decision_text")
ARTIFACTS$decisions$t3b_overall <- tier3b_overall

```

<br />

```{r tier3b-summary, results='asis'}
cat(t3b_summary_text)

```

```{r tier-3b-decision-table, results='asis'}
cat(tier3b_decision_text)

```

<br />

**Tier 3B Overall Decision:** `r tier3b_overall`

<br />

## Tier 3C: Temporal Trends

Temporal trends in `r params$species` for `r contaminant_label` were assessed using linear models across all years of data. An initial model tested for a significant interaction between fish length and sampling year. If an interaction was detected, trends were evaluated separately for the three most frequently sampled length bins. Otherwise, annual mean contaminant concentrations were used. Trends help support Tier 3C by assessing if contaminant levels have declined, remained stable, or increased over time. A half-life for `r contaminant_label` concentrations in `r params$species` was predicted for the AOC and reference sites using a generalized additive mixed model (GAMM). The log values of `r contaminant_label` concentration were estimated using year as a linear predictor by region, with smooth effects controlling for fish length. Sampling sites were used as random effects to account for repeated sampling. The slope (on log scale) represents the annual rate of change in `r contaminant_label` concentration, defined as k = -slope.


```{r temp-trends-setup, message = FALSE}
t3c_data <- filtered_data %>%
  filter(Sample.Year >= 1975, region == "AOC") %>%
  mutate(
      Value = if_else(Value <= 0, 0.001, Value) # GAM breaks on values <= 0, this ensures it works when below detection limits
    )

```

```{r tier-3c-model}
# 
# Fit GAMM
g_gamm <- gamm(
  log(Value) ~ Sample.Year + s(Length, k = 6),
  random = list(Waterbody.Code = ~ 1),
  data = t3c_data
)

t3c_gam <- g_gamm$gam
t3c_lme <- g_gamm$lme


# Helper to get estimate/SE for a linear combo of coefficients
lc_est <- function(coefs, vc, pick) {
  a <- numeric(length(coefs)); names(a) <- names(coefs)
  a[names(pick)] <- pick
  est <- sum(a * coefs)
  se  <- sqrt(as.numeric(t(a) %*% vc %*% a))
  list(est = est, se = se)
}

co <- coef(t3c_gam)
vc <- vcov(t3c_gam)

# Compute slope and half-life
pick <- c(`Sample.Year` = 1)
est <- lc_est(co, vc, pick)
slope <- est$est
slope_se <- est$se

k <- -slope
ln2 <- log(2)

# Calculate half-life or doubling time
t12 <- ifelse(k > 0, ln2 / k, NA_real_)               # declining trend
t2  <- ifelse(k <= 0, ln2 / abs(k), NA_real_)         # increasing trend

# SE via delta method
dtds <- ln2 / (slope^2)
t12_se <- ifelse(!is.na(t12), abs(dtds) * slope_se, NA_real_)
t2_se  <- ifelse(!is.na(t2),  abs(dtds) * slope_se, NA_real_)

# Build single-row tibble for the AOC
hl_tab <- tibble(
  slope_log_per_year = slope,
  slope_se = slope_se,
  k = k,
  half_life_years = t12,
  half_life_lwr = ifelse(!is.na(t12), t12 - 1.96 * t12_se, NA_real_),
  half_life_upr = ifelse(!is.na(t12), t12 + 1.96 * t12_se, NA_real_),
  doubling_time_years = t2,
  doubling_lwr = ifelse(!is.na(t2), t2 - 1.96 * t2_se, NA_real_),
  doubling_upr = ifelse(!is.na(t2), t2 + 1.96 * t2_se, NA_real_)
)

```

```{r half-life-time-to-unrestricted}

# Round to nearest 5 cm
round5 <- function(x) round(x / 5) * 5

# AOC-only years-to-threshold with 95% CI and simplified outputs
compute_years_to_threshold_AOC <- function(
  t3c_gam, hl_tab, thresholds_df, restrict_threshold, t3c_data,
  contaminant_label = "Contaminant"
) {
  # Representative length (upper quartile, rounded to 5 cm)
  quantile_len <- round5(quantile(t3c_data$Length, na.rm = TRUE, probs=0.75))
  
  # Baseline (anchor to most recent observed year in AOC data)
  year_anchor <- max(t3c_data$Sample.Year, na.rm = TRUE)
  
  # Target concentration for the desired meals/month
  target_conc <- thresholds_df %>%
    filter(meals == restrict_threshold) %>%
    pull(conc)
  if (length(target_conc) == 0) stop("restrict_threshold not found in thresholds_df")
  
  # Predict baseline log-concentration at (quantile_len, year_anchor)
  newdata <- tibble(
    Sample.Year = year_anchor,
    Length = quantile_len
  )
  pred <- predict(t3c_gam, newdata = newdata, type = "link", se.fit = TRUE)
  pred_log <- as.numeric(pred$fit)         # log(c0)
  se_log   <- as.numeric(pred$se.fit)      # SE on log-scale
  pred_c0  <- exp(pred_log)                # c0 (current predicted conc)
  
  # Grab k and slope SE (hl_tab is a single AOC row)
  if (nrow(hl_tab) != 1) stop("hl_tab should contain a single AOC row.")
  k        <- hl_tab$k[1]                  # = -slope
  slope_se <- hl_tab$slope_se[1]           # SE of slope (same as SE of k)
  
  # Compute years to threshold and 95% CI (delta method)
  log_target <- log(target_conc)
  years_to_target <- dplyr::case_when(
    pred_c0 <= target_conc ~ 0,
    is.finite(k) & k > 0   ~ (pred_log - log_target) / k,
    TRUE                   ~ NA_real_
  )
  
  se_t <- if (is.finite(years_to_target) && !is.na(years_to_target) && k > 0) {
    # dt/d(log c0) = 1/k ; dt/dk = -(log(c0)-log(c*))/k^2 = -t/k
    term1 <- (1 / k)^2 * (se_log^2)
    term2 <- (years_to_target / k)^2 * (slope_se^2)
    sqrt(term1 + term2)
  } else {
    NA_real_
  }
  lwr <- if (is.na(se_t)) NA_real_ else years_to_target - 1.96 * se_t
  upr <- if (is.na(se_t)) NA_real_ else years_to_target + 1.96 * se_t
  
  # Outcome logic (AOC-only)
  outcome <- dplyr::case_when(
    !is.na(years_to_target) && years_to_target == 0 ~ "Supportive",
    is.na(years_to_target) | k <= 0                 ~ "Unsupportive",
    years_to_target > 10                             ~ "Unsupportive",
    TRUE                                             ~ "Supportive"
  )
  
  # ---- Simplified TIDY TABLE  ----
  results_tbl <- tibble(
    quantile_len_cm   = quantile_len,
    predicted_conc  = pred_c0,
    target_conc     = target_conc,
    years_to_target = years_to_target,
    years_lwr       = lwr,
    years_upr       = upr,
    outcome         = outcome
  )
  
  # ---- Pretty KABLE for reporting ----
  results_kable <- results_tbl %>%
    mutate(
      predicted_conc  = sprintf("%.2f", predicted_conc),
      target_conc     = sprintf("%.2f", target_conc),
      years_to_target = ifelse(is.na(years_to_target), "—", sprintf("%.1f", years_to_target)),
      ci_95 = dplyr::case_when(
        is.na(years_lwr) | is.na(years_upr) ~ "—",
        TRUE                                ~ sprintf("%.1f–%.1f", years_lwr, years_upr)
      )
    ) %>%
    select(
      `Upper Quantile Length (cm)` = quantile_len_cm,
      `Predicted Conc. (Current Year)`     = predicted_conc,
      `Unrestrictive Conc.`        = target_conc,
      `Years to Threshold` = years_to_target,
      `95% CI`             = ci_95,
      `Outcome`            = outcome
    ) %>%
    kable(
      format  = "html",
      caption = glue("Time to reach an unrestrictive advisory of {restrict_threshold} meals/month for upper-quartile-length fish")
    ) %>%
    kable_styling(full_width = FALSE, position = "center")
  
  # ---- Human-friendly METRIC TEXT (short, one-liner) ----
  metric_text <- dplyr::case_when(
    is.na(years_to_target) ~ glue("Trend does not indicate a decline; target not reachable under current trend."),
    years_to_target == 0   ~ glue("Already unrestrictive for upper-quartile-length ({quantile_len} cm) fish."),
    TRUE                   ~ {
      if (!is.na(lwr) && !is.na(upr)) {
        glue("Approximately {round(years_to_target, 1)} years (95% CI {round(lwr, 1)}–{round(upr, 1)}) to reach an unrestrictive meal allowance for upper-quartile-length ({quantile_len} cm) fish.")
      } else {
        glue("Approximately {round(years_to_target, 1)} years to reach an unrestrictive meal allowance for upper-quartile-length ({quantile_len} cm) fish.")
      }
    }
  )
  
  # ---- REPORT TEXT ----
  report_text <- dplyr::case_when(
    is.na(years_to_target) ~ glue(
      "For upper-quartile length ({quantile_len} cm) fish, the current trend does not indicate declining {tolower(contaminant_label)} concentrations. ",
      "Under the present trend, reaching an unrestrictive monthly meal allowance of {restrict_threshold} meals/month for the Sensitive population is not expected."
    ),
    years_to_target == 0 ~ glue(
      "For upper-quartile-length ({quantile_len} cm) fish, predicted concentrations are already within the unrestrictive advisory range for the Sensitive population (≥ {restrict_threshold} meals/month)."
    ),
    TRUE ~ {
      if (!is.na(lwr) && !is.na(upr)) {
        glue(
          "For upper-quartile length ({quantile_len} cm) fish, it is predicted to take about {round(years_to_target, 1)} years ",
          "(95% CI {round(lwr, 1)}–{round(upr, 1)}) to reach an unrestrictive monthly meal allowance of {restrict_threshold} meals/month for the Sensitive population."
        )
      } else {
        glue(
          "For upper-quartile length ({quantile_len} cm) fish, it is predicted to take about {round(years_to_target, 1)} years ",
          "to reach an unrestrictive monthly meal allowance of {restrict_threshold} meals/month for the Sensitive population."
        )
      }
    }
  )
  
  # Outcome string for your tier summary row
  tier3c_overall <- outcome
  
  # Tier 3c summary tibble (mirrors your Tier 3A/3B pattern)
  tier3c_summary_tibble <- tibble(
    Tier = "Tier 3C",
    Description = "Temporal Trends",
    Metric = metric_text,
    Outcome = tier3c_overall
  )
  
  list(
    results_tbl           = results_tbl,     # programmatic table (simplified)
    results_kable         = results_kable,   # report table
    metric_text           = metric_text,     # short one-liner
    report_text           = report_text,     # longer prose (tier-style)
    tier3c_summary_tibble = tier3c_summary_tibble
  )
}


```


```{r Tier-3C-summary-text}

# Compute AOC-only years-to-threshold (uses your revised function signature)
aoc_out <- compute_years_to_threshold_AOC(
  t3c_gam = t3c_gam,
  hl_tab = hl_tab,
  thresholds_df = thresholds_df,
  restrict_threshold = params$restrict_threshold,
  t3c_data = t3c_data,
  contaminant_label = contaminant_label
)

# Show the table
aoc_out$results_kable

# Save artifacts
ARTIFACTS$tables$t3c_results <- save_df(aoc_out$results_tbl, "t3c_results")
ARTIFACTS$html$t3c_results   <- save_kable_html(aoc_out$results_kable, "t3c_results")
ARTIFACTS$text$t3c_report    <- save_text(aoc_out$report_text, "t3c_report_text")

# Outcome (pull from results_tbl since function no longer returns $outcome directly)
ARTIFACTS$decisions$t3c_overall <- aoc_out$results_tbl$outcome[1]

```
```{r tier3c-figure}
# ---- Tier 3c: Future trend visualization (AOC only) ----


# Helper to round to nearest 5 cm (matches your table text)
round5 <- function(x) round(x / 5) * 5

plot_t3c_future <- function(
  t3c_gam,
  t3c_data,
  thresholds_df,
  restrict_threshold,
  contaminant_label = contaminant_label,
  unit = unit_label,
  horizon_years = 20,   # base horizon
  pad_years = 5,        # extra years past the threshold marker
  aoc_out = NULL
) {
  quantile_len <- round5(quantile(t3c_data$Length, na.rm = TRUE, probs = 0.75))
  min_year   <- min(t3c_data$Sample.Year, na.rm = TRUE)
  max_year   <- max(t3c_data$Sample.Year, na.rm = TRUE)
  year_anchor <- max_year

  # If we know years_to_target, make sure we extend at least that far
  vt_year <- NA_real_
  if (!is.null(aoc_out) && is.data.frame(aoc_out$results_tbl)) {
    ytt <- aoc_out$results_tbl$years_to_target[1]
    if (is.finite(ytt) && !is.na(ytt)) {
      vt_year <- year_anchor + ytt
    } else {
      vt_year <- year_anchor + horizon_years
    }
  }

  # End year = max(fixed horizon, threshold year + padding)
  end_year <- max(
    max_year + horizon_years,
    if (is.finite(vt_year)) vt_year + pad_years else -Inf
  )

  proj_years <- seq(min_year, ceiling(end_year), by = 1)

  newdata <- tibble(Sample.Year = proj_years, Length = quantile_len)
  pr <- predict(t3c_gam, newdata = newdata, type = "link", se.fit = TRUE)
  pred_df <- newdata %>%
    mutate(
      fit_log = as.numeric(pr$fit),
      se_log  = as.numeric(pr$se.fit),
      conc    = exp(fit_log),
      lwr     = exp(fit_log - 1.96 * se_log),
      upr     = exp(fit_log + 1.96 * se_log),
      period  = ifelse(Sample.Year <= max_year, "Observed", "Projected")
    )

  target_conc <- thresholds_df %>% filter(meals == restrict_threshold) %>% pull(conc)
  if (length(target_conc) != 1) stop("Could not identify a single target concentration from thresholds_df.")

  vt_df <- if (is.finite(vt_year)) tibble(vline_x = vt_year) else NULL
  thr_lab <- paste0("Unrestrictive threshold (", restrict_threshold, " meals/mo)")

  ggplot(pred_df, aes(x = Sample.Year, y = conc)) +
    geom_ribbon(aes(ymin = lwr, ymax = upr), alpha = 0.15, linewidth = 0) +
    geom_line(data = dplyr::filter(pred_df, period == "Observed"),  linewidth = 1) +
    geom_line(data = dplyr::filter(pred_df, period == "Projected"), linewidth = 1, linetype = "22") +
    geom_hline(yintercept = target_conc, linetype = "dotted") +
    annotate("text",
             x = min_year + 0.02 * (max(proj_years) - min_year),
             y = target_conc, label = thr_lab, vjust = -0.6, hjust = 0, size = 3.4) +
    scale_y_continuous(labels = scales::label_number(accuracy = 0.1)) +
    scale_x_continuous(breaks = seq((round(min_year/5)*5),(round(vt_year/5)*5),10))+
    labs(
      x = "Year",
      y = paste0(contaminant_label, " concentration", " (", unit_label, ")"),
      subtitle = paste0("Representative fish length = ", quantile_len, " cm"),
      caption = "Solid = observed; Dashed = projection; Dotted = unrestrictive threshold"
    ) +
    theme_classic(base_size = 12) +
    theme(plot.title = element_text(face = "bold"),
          legend.position = "none")
}


# ---- Example call ----
p_t3c <- plot_t3c_future(
  t3c_gam = t3c_gam,
  t3c_data = t3c_data,
  thresholds_df = thresholds_df,
  restrict_threshold = params$restrict_threshold,
  contaminant_label = contaminant_label,
  unit = unit_label,
  horizon_years = 20,
  pad_years = 0,
  aoc_out = aoc_out
)

p_t3c

ARTIFACTS$figs$t3c_projected <- save_gg(p_t3c, "t3c_projected")
```
A model test was performed to estimate the time required to reach a nonrestrictive monthly meal allowance for the upper quantile fish length (rounded to the nearest 5 cm) for sensitive populations. A timeline of 10 years or fewer is considered supportive of delisting, whereas a timeline of 10 or more years is unsupportive.

`r aoc_out$report_text` Therefore, Tier 3C is considered `r aoc_out$outcome`.

**Tier 3C Overall Decision:** `r aoc_out$outcome`
```{r new-t3c-method}
# Round to nearest 5 cm (reuse your existing helper if already defined)
round5 <- function(x) round(x / 5) * 5

# Vectorized AOC-only years-to-threshold over multiple representative lengths
compute_years_to_threshold_lengths <- function(
  t3c_gam,
  hl_tab,
  thresholds_df,
  restrict_threshold,
  t3c_data,
  rep_lengths_cm,
  contaminant_label = "Contaminant"
) {
  # Anchor to most recent observed year in AOC data
  year_anchor <- max(t3c_data$Sample.Year, na.rm = TRUE)
  
  # Target concentration for desired meals/month
  target_conc <- thresholds_df %>%
    dplyr::filter(meals == restrict_threshold) %>%
    dplyr::pull(conc)
  if (length(target_conc) == 0) stop("restrict_threshold not found in thresholds_df")
  if (length(target_conc) > 1) stop("thresholds_df must yield a single target_conc.")
  
  # Grab k and slope SE (AOC-only)
  if (nrow(hl_tab) != 1) stop("hl_tab should contain a single AOC row.")
  k        <- hl_tab$k[1]          # = -slope
  slope_se <- hl_tab$slope_se[1]   # SE of slope (same as SE of k)
  
  log_target <- log(target_conc)
  
  purrr::map_dfr(rep_lengths_cm, function(len_cm) {
    newdata <- tibble::tibble(
      Sample.Year = year_anchor,
      Length      = len_cm
    )
    pred <- predict(t3c_gam, newdata = newdata, type = "link", se.fit = TRUE)
    pred_log <- as.numeric(pred$fit)
    se_log   <- as.numeric(pred$se.fit)
    pred_c0  <- exp(pred_log)
    
    years_to_target <- dplyr::case_when(
      pred_c0 <= target_conc ~ 0,
      is.finite(k) & k > 0   ~ (pred_log - log_target) / k,
      TRUE                   ~ NA_real_
    )
    
    se_t <- if (is.finite(years_to_target) && !is.na(years_to_target) && k > 0) {
      # dt/d(log c0) = 1/k; dt/dk = -(log(c0) - log(c*))/k^2 = -t/k
      term1 <- (1 / k)^2        * (se_log^2)
      term2 <- (years_to_target / k)^2 * (slope_se^2)
      sqrt(term1 + term2)
    } else {
      NA_real_
    }
    
    lwr <- if (is.na(se_t)) NA_real_ else years_to_target - 1.96 * se_t
    upr <- if (is.na(se_t)) NA_real_ else years_to_target + 1.96 * se_t
    
    outcome <- dplyr::case_when(
      !is.na(years_to_target) && years_to_target == 0 ~ "Supportive",
      is.na(years_to_target) | k <= 0                 ~ "Unsupportive",
      years_to_target > 10                             ~ "Unsupportive",
      TRUE                                             ~ "Supportive"
    )
    
    tibble::tibble(
      length_cm       = len_cm,
      predicted_conc  = pred_c0,
      target_conc     = target_conc,
      years_to_target = years_to_target,
      years_lwr       = lwr,
      years_upr       = upr,
      outcome         = outcome
    )
  })
}





# Build a tibble of representative lengths (quartiles by default)
get_rep_lengths_quartiles <- function(
  t3c_data,
  probs = c(0.25, 0.5, 0.75),
  round_fun = round5
) {
  qs <- stats::quantile(t3c_data$Length, probs = probs, na.rm = TRUE)
  tibble::tibble(
    role      = c("Lower quartile", "Median", "Upper quartile")[seq_along(qs)],
    prob      = probs,
    length_cm = round_fun(as.numeric(qs))
  )
}

# Example alternative for survey-based sizes
# rep_info <- tibble::tibble(
#   role      = c("Survey P50", "Survey P75"),
#   prob      = c(NA_real_, NA_real_),
#   length_cm = c(45, 55)
# )





# Choose your configuration:
# 1) All three quartiles:
rep_info <- get_rep_lengths_quartiles(t3c_data)

# 2) Only upper quartile:
# rep_info <- get_rep_lengths_quartiles(t3c_data) %>%
#   dplyr::filter(role == "Upper quartile")

# 3) Survey-based:
# rep_info <- tibble::tibble(
#   role      = "Survey representative size",
#   prob      = NA_real_,
#   length_cm = 52  # for example
# )

t3c_sensitivity_tbl <- compute_years_to_threshold_lengths(
  t3c_gam          = t3c_gam,
  hl_tab           = hl_tab,
  thresholds_df    = thresholds_df,
  restrict_threshold = new_thr,
  t3c_data         = t3c_data,
  rep_lengths_cm   = rep_info$length_cm,
  contaminant_label = contaminant_label
) %>%
  dplyr::left_join(rep_info, by = "length_cm") %>%
  dplyr::relocate(role, .before = length_cm)





t3c_sensitivity_kable <- t3c_sensitivity_tbl %>%
  dplyr::mutate(
    predicted_conc  = sprintf("%.2f", predicted_conc),
    target_conc     = sprintf("%.2f", target_conc),
    years_to_target = dplyr::if_else(
      is.na(years_to_target),
      "—",
      sprintf("%.1f", years_to_target)
    ),
    ci_95 = dplyr::case_when(
      is.na(years_lwr) | is.na(years_upr) ~ "—",
      TRUE ~ sprintf("%.1f–%.1f", years_lwr, years_upr)
    )
  ) %>%
  dplyr::select(
    `Representative group`     = role,
    `Length (cm)`              = length_cm,
    `Predicted Conc. (Current Year)` = predicted_conc,
    `Unrestrictive Conc.`      = target_conc,
    `Years to Threshold`       = years_to_target,
    `95% CI`                   = ci_95,
    `Outcome`                  = outcome
  ) %>%
  kableExtra::kable(
    format  = "html",
    caption = glue::glue(
      "Time to reach an unrestrictive advisory of {params$restrict_threshold} ",
      "meals/month for different representative fish lengths"
    )
  ) %>%
  kableExtra::kable_styling(full_width = FALSE, position = "center")

# Optionally cache
ARTIFACTS$tables$t3c_sensitivity <- save_df(t3c_sensitivity_tbl, "t3c_sensitivity")
ARTIFACTS$html$t3c_sensitivity   <- save_kable_html(t3c_sensitivity_kable, "t3c_sensitivity")

t3c_sensitivity_kable






plot_t3c_future_multi <- function(
  t3c_gam,
  t3c_data,
  thresholds_df,
  restrict_threshold,
  rep_info,              # tibble with at least `role` and `length_cm`
  contaminant_label = contaminant_label,
  unit = unit_label,
  horizon_years = 20,
  primary_role = "Median",   # which length gets full opacity & CI ribbon
  shade = c("primary", "all", "none"),  # CI shading behavior
  cut_info = NULL,           # tibble with `role` and `years_to_target`
  cut_role = "Upper quartile",
  pad_years = 0              # extend a bit beyond intersect, if desired
) {
  shade <- match.arg(shade)
  
  # Custom colour palette for representative lengths
colour_vals <- c(
  "Lower quartile" = "#696969",
  "Median"         = "black",
  "Upper quartile" = "#ADADAD"
)


  # ---- Basic time range ----
  min_year   <- min(t3c_data$Sample.Year, na.rm = TRUE)
  max_year   <- max(t3c_data$Sample.Year, na.rm = TRUE)
  anchor_year <- max_year   # same anchor you use in the years-to-threshold calc

  # ---- Compute cut_year from most restrictive length (if available) ----
  cut_year <- NA_real_
  if (!is.null(cut_info)) {
    ytt <- cut_info %>%
      dplyr::filter(.data$role == cut_role) %>%
      dplyr::pull(.data$years_to_target)

    if (length(ytt) == 1 && is.finite(ytt) && !is.na(ytt) && ytt > 0) {
      cut_year <- anchor_year + ytt
    }
  }

  # If no valid cut_year, just use horizon_years beyond observed
  end_year <- if (is.finite(cut_year)) {
    cut_year + pad_years
  } else {
    max_year + horizon_years
  }

  proj_years <- seq(min_year, ceiling(end_year), by = 1)

  # ---- Target concentration for unrestrictive meals/month ----
  target_conc <- thresholds_df %>%
    dplyr::filter(.data$meals == restrict_threshold) %>%
    dplyr::pull(.data$conc)

  if (length(target_conc) != 1) {
    stop("Could not identify a single target concentration from thresholds_df.")
  }

  # ---- Predict for each representative length ----
  pred_df <- purrr::map_dfr(seq_len(nrow(rep_info)), function(i) {
    len_cm <- rep_info$length_cm[i]
    role   <- rep_info$role[i]

    newdata <- tibble::tibble(
      Sample.Year = proj_years,
      Length      = len_cm,
      role        = role
    )

    pr <- predict(t3c_gam, newdata = newdata, type = "link", se.fit = TRUE)

    newdata %>%
      dplyr::mutate(
        fit_log = as.numeric(pr$fit),
        se_log  = as.numeric(pr$se.fit),
        conc    = exp(fit_log),
        lwr     = exp(fit_log - 1.96 * se_log),
        upr     = exp(fit_log + 1.96 * se_log),
        period  = dplyr::if_else(.data$Sample.Year <= max_year, "Observed", "Projected")
      )
  })

  # ---- Alpha mapping: primary fully opaque, others lighter ----
  roles <- unique(pred_df$role)
  alpha_vals <- rep(0.5, length(roles))
  names(alpha_vals) <- roles
  if (primary_role %in% roles) alpha_vals[primary_role] <- 1

  # ---- Base plot ----
  p <- ggplot(pred_df, aes(x = Sample.Year, y = conc, colour = role)) +
     scale_colour_manual(values = colour_vals)


  # ---- Optional CI shading ----
  if (shade != "none") {
    if (shade == "primary") {
      ribbon_data <- dplyr::filter(pred_df, role == primary_role)
    } else {  # "all"
      ribbon_data <- pred_df
    }

    p <- p +
      geom_ribbon(
        data = ribbon_data,
        aes(
          x    = Sample.Year,
          ymin = lwr,
          ymax = upr,
          group = role
        ),
        inherit.aes = FALSE,
        alpha       = 0.12,
        linewidth   = 0
      )
  }

  # ---- Lines for each representative length ----
  p <- p +
    geom_line(
      aes(alpha = role, linetype = period),
      linewidth = 1
    ) +
    geom_hline(yintercept = target_conc, linetype = "dotted")

  thr_lab <- paste0("Unrestrictive threshold (", restrict_threshold, " meals/mo)")

  p <- p +
    annotate(
      "text",
      x      = min_year + 0.02 * (max(proj_years) - min_year),
      y      = target_conc,
      label  = thr_lab,
      vjust  = -0.6,
      hjust  = 0,
      size   = 3.4
    ) +
    scale_alpha_manual(values = alpha_vals, guide = "none") +
    scale_linetype_manual(values = c(Observed = "solid", Projected = "22")) +
    scale_y_continuous(labels = scales::label_number(accuracy = 0.1)) +
    scale_x_continuous(breaks = pretty(proj_years, n = 8)) +
    labs(
      x        = "Year",
      y        = paste0(contaminant_label, " concentration (", unit, ")"),
      colour   = "Representative length",
      linetype = "",
      caption = paste(
        "Representative fish lengths:",
        paste0(rep_info$role, " (", rep_info$length_cm, " cm)", collapse = "; ")
      )
    ) +
    theme_classic(base_size = 12) +
    theme(
      plot.title      = element_text(face = "bold"),
      legend.position = "right"
    )

  p

  }


# Same data from table
cut_info <- t3c_sensitivity_tbl %>%
  dplyr::select(role, years_to_target)

# Use the same rep_info as in the table
rep_info <- get_rep_lengths_quartiles(t3c_data)
# or any custom tibble with `role` + `length_cm`

p_t3c_multi <- plot_t3c_future_multi(
  t3c_gam            = t3c_gam,
  t3c_data           = t3c_data,
  thresholds_df      = thresholds_df,
  restrict_threshold = params$restrict_threshold,
  rep_info           = rep_info,          # e.g. quartiles tibble
  contaminant_label  = contaminant_label,
  unit               = unit_label,
  horizon_years      = 20,                # used if no valid cut_year
  primary_role       = "Median",
  shade              = "all",
  cut_info           = cut_info,
  cut_role           = "Upper quartile",  # "most restrictive"
  pad_years          = 0                  # or 1–2 if you want a tiny bit past
)


p_t3c_multi

ARTIFACTS$figs$t3c_projected_multi <- save_gg(p_t3c_multi, "t3c_projected_multi")


```
# Final summary

```{r}
tier3_summary_table <- bind_rows(
  tier3a_summary_tibble,
  tier3b_summary_tibble,
  aoc_out$tier3c_summary_tibble
)

knitr::kable(tier3_summary_table, caption = "Summary of Tier 3 Evidence Lines and Outcomes")

ARTIFACTS$tables$summary <- save_df(tier3_summary_table, "tier3_summary_table")
ARTIFACTS$summary <- tier3_summary_table

# Write a compact index.rds pointing to everything you saved
idx_file <- file.path(out_dir, "index.rds")
# … after populating ARTIFACTS …
saveRDS(ARTIFACTS, file.path(out_dir_abs, "index.rds"))
writeLines(c(
  glue("species: {params$species}"),
  glue("contaminant: {params$contaminant}"),
  glue("output_dir_rel: {out_dir_rel}"),       # useful for debugging
  glue("index_rds_rel: index.rds")
), file.path(out_dir_abs, "INDEX.txt"))


```


```{r t3c-smooth, eval = FALSE}
library(mgcv)

m_timevary <- gam(
  log(Value) ~ 
    s(Sample.Year, k = 8) +         # flexible, but modest k; increase if needed
    s(Length, k = 6) + 
    s(Waterbody.Code, bs = "re"),   # random intercept = GAMM-like
  data = t3c_data,
  method = "REML"
)
summary(m_timevary)

# Instantaneous rate k(t) and half-life over time
# Rate on log scale is negative derivative of time smooth
library(gratia)

der <- derivatives(m_timevary, select = "s(Sample.Year)", order = 1)
# k(t) = - d/dt E[log(Value)]
der$k_t <- -der$.derivative
der$k_se <- der$.se
der$half_life <- ifelse(der$k_t > 0, log(2) / der$k_t, NA_real_)

round5 <- function(x) round(x/5)*5
median_len <- round5(median(t3c_data$Length, na.rm = TRUE))

target_conc <- thresholds_df$conc[match(params$restrict_threshold, thresholds_df$meals)]
eta_minus_logtarget <- function(y) {
  predict(m_timevary,
          newdata = data.frame(Sample.Year = y, Length = median_len),
          type = "link") - log(target_conc)
}

anchor <- max(t3c_data$Sample.Year, na.rm = TRUE)
# extend far enough; adjust upper bound if needed
root <- try(uniroot(eta_minus_logtarget, lower = anchor, upper = anchor + 150), silent = TRUE)

years_to_target <- if (inherits(root, "try-error")) NA_real_ else (root$root - anchor)




library(dplyr)
library(tidyr)
library(ggplot2)
library(scales)

round5 <- function(x) round(x/5)*5

plot_t3c_future_timevary <- function(
  m_timevary, t3c_data, thresholds_df, restrict_threshold,
  contaminant_label = "Contaminant",
  horizon_years = 20, pad_years = 5,
  aoc_out = NULL  # optional: to extend to threshold year
) {
  median_len <- round5(median(t3c_data$Length, na.rm = TRUE))
  min_year   <- min(t3c_data$Sample.Year, na.rm = TRUE)
  max_year   <- max(t3c_data$Sample.Year, na.rm = TRUE)
  anchor     <- max_year

  # Compute threshold year from aoc_out if available
  vt_year <- NA_real_
  if (!is.null(aoc_out) && is.data.frame(aoc_out$results_tbl)) {
    ytt <- aoc_out$results_tbl$years_to_target[1]
    if (is.finite(ytt) && !is.na(ytt)) vt_year <- anchor + ytt
  }

  end_year <- max(max_year + horizon_years, if (is.finite(vt_year)) vt_year + pad_years else -Inf)
  proj_years <- seq(min_year, ceiling(end_year), by = 1)

  newdata <- tibble(Sample.Year = proj_years, Length = median_len)
  pr <- predict(m_timevary, newdata = newdata, type = "link", se.fit = TRUE)
  pred_df <- newdata %>%
    mutate(
      fit_log = as.numeric(pr$fit),
      se_log  = as.numeric(pr$se.fit),
      conc    = exp(fit_log),
      lwr     = exp(fit_log - 1.96 * se_log),
      upr     = exp(fit_log + 1.96 * se_log),
      period  = ifelse(Sample.Year <= max_year, "Observed", "Projected")
    )

  target_conc <- thresholds_df %>% filter(meals == restrict_threshold) %>% pull(conc)
  stopifnot(length(target_conc) == 1)

  vt_df <- if (is.finite(vt_year)) tibble(vline_x = vt_year) else NULL
  thr_lab <- paste0("Unrestrictive threshold (", restrict_threshold, " meals/mo)")

  ggplot(pred_df, aes(Sample.Year, conc)) +
    geom_ribbon(aes(ymin = lwr, ymax = upr), alpha = 0.15, linewidth = 0) +
    geom_line(data = filter(pred_df, period == "Observed"),  linewidth = 1) +
    geom_line(data = filter(pred_df, period == "Projected"), linewidth = 1, linetype = "22") +
    geom_hline(yintercept = target_conc, linetype = "dotted") +
    { if (!is.null(vt_df)) geom_vline(data = vt_df, aes(xintercept = vline_x),
                                      linetype = "dotdash", linewidth = 0.6, colour = "grey30") } +
    annotate("text",
             x = min_year + 0.02 * (max(proj_years) - min_year),
             y = target_conc, label = thr_lab, vjust = -0.6, hjust = 0, size = 3.4) +
    { if (!is.null(vt_df)) annotate("text",
                                    x = vt_df$vline_x, y = max(pred_df$conc, na.rm = TRUE),
                                    label = "Reach threshold (est.)",
                                    vjust = 1.2, hjust = 0.5, size = 3.2, colour = "grey30") } +
    scale_y_continuous(labels = label_number(accuracy = 0.1)) +
    labs(
      x = "Year",
      y = paste0(contaminant_label, " concentration"),
      title = paste0("Tier 3c: Projected ", contaminant_label, " (AOC, time-varying trend)"),
      subtitle = paste0("Representative fish length = ", median_len, " cm; ribbons = 95% CI"),
      caption = "Solid = observed; Dashed = projection; Dotted = unrestrictive threshold"
    ) +
    theme_minimal(base_size = 12) +
    theme(plot.title = element_text(face = "bold"),
          legend.position = "none")
}

# Example call
p_t3c_tv <- plot_t3c_future_timevary(
  m_timevary = m_timevary,
  t3c_data   = t3c_data,
  thresholds_df = thresholds_df,
  restrict_threshold = new_thr,
  contaminant_label = contaminant_label,
  horizon_years = 20,
  pad_years = 5,
  aoc_out = aoc_out
)
p_t3c_tv



library(gratia)

# Smooth plot
draw(m_timevary, select = "s(Sample.Year)")

# Derivative = d/dt log(Value); rate k(t) = -derivative
der <- derivatives(m_timevary, term = "s(Sample.Year)", order = 1)
der$k_t <- -der$derivative

ggplot(der, aes(data, k_t)) +
  geom_hline(yintercept = 0, linetype = "dotted") +
  geom_ribbon(aes(ymin = lower, ymax = upper), alpha = 0.15) +
  geom_line() +
  labs(x = "Year", y = "Instantaneous decline rate k(t)",
       title = "Time-varying rate of decline (k(t))") +
  theme_minimal()

```


